\chapter{NNLL resummation in $Z$+jet events with the ARES method}
\label{ch:hhZjet}

\section{Introduction}
\label{sec:intro}

To date, no general method has been established for performing
NNLL-accurate resummation of jet observables in hadronic collisions.
Considerable effort is currently devoted to extending parton showers
to this accuracy, but the current state of the art remains at NLL,
demonstrated by \textsc{Sherpa}~\cite{Herren:2022jej,Sherpa:2024mfk}
and \textsc{PanScales}~\cite{vanBeekveld:2022zhl,vanBeekveld:2022ukn,
vanBeekveld:2023chs}. Within Soft-Collinear Effective Theory (SCET),
resummation has been achieved up to N$^3$LL accuracy for a number of
hadronic collision observables, but these approaches are
observable-specific, each relying on a dedicated factorisation theorem.
Although progress has recently been made towards a more general
framework~\cite{vanBeekveld:2025zjh}, no formalism currently exists
that applies to a broad class of event shapes.

In this chapter we extend the \textsc{ARES} formalism for
next-to-next-to-leading logarithmic (NNLL) resummation to hadronic
collisions, specifically to the process $pp\to Z+\text{jet}$, where
the $Z$ boson recoils against a jet with high transverse momentum. We
find that the structure of the NNLL resummation is formally identical
to that for three-jet events in $e^+e^-$ annihilation~\cite{Arpino:2019ozn}, with the same
master formula but different ingredients. In particular, initial-state
radiation at NNLL involves the convolution of splitting functions with
parton distribution functions --- a new contribution absent in the
$e^+e^-$ case. 

To illustrate the method, we consider the event-shape observable
one-jettiness, which has been studied extensively and already resummed
to N$^3$LL accuracy~\cite{Alioli:2023rxx}, providing a useful
benchmark against which to validate our results. One-jettiness is also
an additive observable, making it particularly straightforward to treat
within \textsc{ARES}. It therefore serves as an ideal worked example
through which to demonstrate the implementation of the formalism,
rather than as a source of new phenomenological results. We validate
our results analytically against the known resummation of one-jettiness.

Unlike the NLL discussion presented in the previous chapter, we do not
derive the NNLL master formula from first principles, as this is not
the primary objective here. Instead, we focus on its extension to
hadronic collisions, in particular the treatment of initial-state
radiation. The derivation of the general NNLL master formula can be
found in Refs.~\cite{Banfi:2014sua,Banfi:2018mcq,Arpino:2019ozn}, to
which we refer the reader for a detailed discussion.

The remainder of this chapter is organised as follows. We begin by
presenting the \textsc{ARES} NNLL master formula for hadronic
collisions, before deriving the initial-state radiation contributions.
We then apply the formalism to one-jettiness, and close with a
validation of the results against existing calculations in the
literature.





\section{General setup}
\label{sec:setup}

We consider the production of a $Z$ boson and a hard jet in hadron collisions,
\begin{equation}
  \label{eq:Z+jet}
  h_1(P_1)+h_2(P_2) \to Z(q) + \mathrm{jet}(p_J)+\dots\,,
\end{equation}
where $h_1$ and $h_2$ denote the incoming hadrons with momenta $P_1$ and $P_2$, and the dots represents accompanying hadrons. At Born level, the process proceeds through two incoming partons with momenta $p_1=x_1P_1$ and $p_2=x_2P_2$, producing a $Z$ boson with momentum $q$ and a single hard parton with momentum $p_3$. Beyond Born level, $p_3$ will recoil
against additional partons. Its momentum will change to
$\tilde p_3$, and it will be clustered with other partons to form a hard jet $p_J$. 
At Born level, the two coincide, such that $p_J=p_3$.

We consider recursively infrared and collinear (rIRC) safe event
shapes, which we denote generally as functions $V(q_1,\dots,q_n)$ of
hadron momenta $q_1, \dots, q_n$. As an example, we consider
1-jettiness, an observable that is widely studied in SCET and is the
$N=1$ case of the more general $N$-jettiness observable~\cite{Stewart:2010tn}, 
which quantifies how well the final state can be described by $N$ narrow jets.. This is
defined as
\begin{equation}
 \label{eq:T1}
 \mathcal{T}_1\equiv \sum_{i} \min\left\{\frac{2 \bar{p}_1\cdot q_i}{Q_1}, \frac{2 \bar{p}_2\cdot q_i}{Q_2}, \frac{2 p_J \cdot q_i}{Q_3}\right\}\,,
\end{equation}
where $q_i$ is the four-momentum of a \emph{hadron}, and
$Q_1, Q_2, Q_3$ three momentum scales. The two ``incoming'' momenta, 
$\bar p_1$ and $\bar p_2$ are defined as $\bar p_i=E_i n_i$, where
\begin{equation}
	\bar{p}_1^\mu=\frac{1}{2}n_2\cdot (p_J+q)n_1^{\mu}\,,\quad\quad
	\bar{p}_2^\mu=\frac{1}{2}n_1\cdot (p_J+q)n_2^{\mu}\,,\quad\quad n_i^{\mu}=(1,\vec{n}_i)\,,\quad\quad |\vec{n}_i|=1\,.
\end{equation}
\begin{equation}
  \label{eq:ni-incoming}
  n_1 = (1,0,0,1)\,,\qquad n_2 = (1,0,0,-1)\,\qquad E_i= n_i \cdot (p_J+q)\,.
\end{equation}
When $\mathcal{T}_1\to 0$, we are in the ``1-jet region'', in which
the outgoing jet $p_J$ recoils against the $Z$ boson. In that case,
$\bar p_1$ and $\bar p_2$ coincide with the Born momenta $p_1$ and
$p_2$.

We can also consider a normalised version of 1-jettiness,
$\tau_1\equiv \mathcal{T}_1/M$, where $M$ denotes the hard scale of the process. 
A particular choice of $Q_1, Q_2, Q_3$
\begin{equation}
  \label{eq:Qi-inv}
  Q_1 = \frac{2 \bar{p}_1\cdot p_J}{p_{T,J}}\,,  \quad Q_2 = \frac{2 \bar{p}_2\cdot p_J}{p_{T,J}}\,,
  \quad Q_3 = 2 p_{T,J}\,,
\end{equation}
where $p_{T,J}$ is the transverse momentum of the outgoing jet with
respect to the beam, provides a definition of $\mathcal{T}_1$ that is
manifestly invariant under boosts along the beam direction:
\begin{equation}
	\label{eq:T1-binv}
	\mathcal{T}_1\equiv  \sum_{i} k_{t,i}\min\{e^{(\eta_i-\eta_J)}, e^{-(\eta_i-\eta_J)}, \cosh(\eta_{i}-\eta_J)-\cos(\phi_{i}-\phi_J)\}\,,
\end{equation}
where $k_{t,i}$ and $\eta_i$ are the transverse momentum and rapidity
of hadron $q_i$ with respect to the beam, and $\eta_J$ the rapidity of the hard jet. 
To reduce the impact of experimental systematics in the measurement of particle momenta, 
we can set $M=H_T$, with $H_T$ the scalar sum of all hadron transverse
momenta, leading to a dimensionless definition of $\tau_1$, namely
\begin{equation}
	\label{eq:tau1-binv}
	\tau_1\equiv  \sum_{i} \frac{k_{t,i}}{H_T}\min\{e^{(\eta_i-\eta_J)}, e^{-(\eta_i-\eta_J)}, \cosh(\eta_{i}-\eta_J)-\cos(\phi_{i}-\phi_J)\}\,.
\end{equation}

In general, we want the cumulant of an event shape, defined by the constraint that a given
event shape $V(q_1,q_2,\dots,q_n)$ is less than a given value
$v$. This observable is such that the one-jet region corresponds
to $v\to 0$. Additionally, in the one-jet region, we wish \question{to use} Born kinematics completely. 
This can be obtained for instance by measuring the differential cross-section in $\eta_J$, and in the transverse momentum $q_T$ and rapidity $y_Z$ of the $Z$ boson.  In the one-jet limit, the aforementioned
cross-section takes the form
\begin{equation}
  \label{eq:dsigma-def}
  \frac{d\sigma(v)}{dq_T dy_Zd\eta_J} \simeq \sum_{i,j} \int_0^1 \!\! dx_1\, f_i(x_1,\mu_F)\int_0^1 \!\! dx_2\, f_j(x_2,\mu_F)\frac{d\hat \sigma_{ij}}{dq_Tdy_Zd\eta_J} \Sigma_{ij}(\{p_1,p_2,p_3\},v)\,,
\end{equation}
\ABquestion{I would like to refer to the cumulant in Sec.~\ref{sec:problem-spec},
but unsure how they map onto each other.}
In the above equation, we have introduced the partonic cross section
for the production of a $Z$ boson and an extra jet via the partonic
subprocess $ij$, where $i$ represents a quark, an antiquark of any
light flavour ($u,d,c,s,b$, considered massless), or a gluon. The expression in Eq.~\eqref{eq:dsigma-def}
requires the construction of a mapping between the full event, made up
of the three hard partons $\{\tilde p\}\equiv\{\tilde p_1,\tilde p_2,\tilde p_3\}$ and an arbitrary number of soft and/or collinear emissions $\{k_i\}$, and
a Born event with only three hard partons $\{p_1,p_2,p_3\}$. The
effect of multiple soft and/or collinear emissions, and the
corresponding virtual corrections, for fixed Born kinematics \question{do you mean fix here?}. The difference between the approximation on the right-hand side of
eq.~\eqref{eq:dsigma-def} and the full expression for
$ d\sigma(v)/(dq_T dy_Z d\eta_J)$ is a well-defined expansion in powers of the
strong coupling $\alpha_s$, with coefficients vanishins as a power of
$v$, potentially enhanced by logarithms of $v$.

The cumulative distribution $\Sigma_{ij}(\{p_1,p_2,p_3\},v)$ contains
logarithms of $v$ that become large in the near-to-planar region. The
largest of these contributions are double logarithms of the form
$(\alpha_s L^2)^n$, with $L\equiv \ln(1/v)$. Neglecting for the moment
coherence-violating effects, if $V(\{\tilde p\},\{k_i\})$ is rIRC
safe, for each Born phase space point, in the region
$\alpha_s L\sim 1$, $\Sigma_{ij}(\{p_1,p_2,p_3\},v)$ can be
reorganised as follows
\begin{equation}
  \label{eq:log-counting}
  \Sigma_{ij}(\{p_1,p_2,p_3\},v) = e^{L g_1(\alpha_s L)}\left[G_2(\alpha_s L)+ \alpha_s G_3(\alpha_s L)+\dots\right]\,.
\end{equation}
\ABquestion{Shall I refer to how this is different to that in Eq.~\eqref{eq:resummed-cumulant}?}
As illustrated in Table.~\ref{tab:log-accuracy} the function $g_1(\alpha_s L)$ resums all leading logarithmic (LL)
terms, $\alpha_s^n L^{n+1}$, in $\ln\Sigma_{ij}$. For rIRC safe
observables, $g_1$ exponentiates. The function $G_2(\alpha_s L)$
resums next-to-leading logarithmic (NLL) contributions
($\alpha_s^n L^n$) to $\ln\Sigma_{ij}$, and so on. In this work, we
aim at NNLL accuracy, i.e.\ calculating $g_1$, $G_2$ and $G_3$. 

 
\section{General form of NNLL resummation}
\label{sec:NNLL-general}

The main differences between event shape resummation in hadronic
collisions and $e^+e^-$ annihilation are
\begin{itemize}
\item the presence of initial-state radiation;
\item the fact that there is no single natural hard scale $Q$.
\end{itemize}
Nevertheless, as will be made clear in Sec.~\ref{sec:NNLL-initial}, the structure of the NNLL resummation for $\Sigma_{ij}(\{p_1,p_2,p_3\},v)$ is the same as for $e^+e^-$ annihilation, but some of the ingredients appearing in the resummation formula have to be slightly modified. In this section, we present the general structure of the NNLL resummation, and discuss the physical meaning of each ingredient. For each partonic channel $ij$, we have 
\begin{multline}
  \label{eq:Sigma-NNLL}
  \Sigma_{ij}(\{p_1,p_2,p_3\},v)= e^{-R_{\rm NNLL}(v)} \frac{f_i(x_1,v^{\frac{1}{a+b_1}}Q)}{f_i(x_1,\mu_F)} \frac{f_j(x_2,v^{\frac{1}{a+b_2}}Q)}{f_j(x_2,\mu_F)}
  \times \\ \times
  \left(\mathcal{F}^{(ij)}_{\rm NLL}(\lambda)\left(1+\frac{\alpha_s(Q)}{2\pi}H_{ij}^{(1)}(Q,\mu_R)
      +\frac{\alpha_s(Q v^{\frac{1}{a+b_\ell}})}{2\pi}\sum_{\ell}C_{\mathrm{hc},\ell}^{(1)}\right) +\frac{\alpha_s(Q)}{\pi} \delta\mathcal{F}^{(ij)}_{\rm NNLL}(\lambda)\right)\,,
\end{multline}
where $\mu_R,\mu_F$ are the renormalisation and factorisation
scales in the $\overline{\rm MS}$ scheme, and $\beta_0$ and $\lambda$
are defined in Eqs.~\eqref{eq:beta0} and~\eqref{eq:lambda} respectively.
The parameters $a_\ell, b_\ell$ are determined by the general behaviour 
of an IRC safe observable $V$ in the presence of a single soft emission collinear to leg $\ell$
as in Eq.~\eqref{eq:general-sc-V}. Note that for $\ee$ annihilation, the scale $Q$ 
is a typical hard scale of the process under consideration, which by default can be chosen
to be the centre-of-mass-energy of the $\ee$ collision. In hadronic
collisions, on the other hand, this scale is an arbitrary ``resummation' scale,
we use as a reference normalisation scale for transverse momenta. 

The physical meaning of all ingredients in
Eq.~\eqref{eq:Sigma-NNLL} can be recovered from
Refs.~\cite{Arpino:2019ozn,Banfi:2018mcq,Banfi:2004yd}. We give
below a short description of each, highlighting the differences with
respect to the $e^+e^-$ case:

\begin{itemize}
\item The NNLL radiator $R_{\rm NNLL}(v)$ includes all virtual
  corrections of soft and/or collinear origin. These are divergent,
  but their divergences cancel with those of real radiation. In
  order to cancel the singularities of virtual corrections analytically, the
  radiator also includes also a subset of real corrections for suitably
  defined ``unresolved'' emissions. The procedure to define those
  emissions is not unique. The procedure to define these emissions is
  not unique. A proposal adequate for three emitting legs is given in
  Ref.~\cite{Arpino:2019ozn}, leading to a master formula for
  $R_{\rm NNLL}(v)$ whose explicit expression depends only on the
  single-emission parameters $a$, $b_\ell$, $d_\ell$, and
  $g_\ell(\phi)$. The same master formula applies in hadronic
  collisions, with $Q$ reinterpreted as the resummation scale rather
  than the centre-of-mass energy. For completeness, we illustrate how
  the NNLL radiator extends from the NLL result in
  Sec.~\ref{sec:radiator}. It is worth noting that  
  with the knowledge we have acquired so far, the radiator 
  could be computed for a generic observable at N$^3$LL accuracy. 
  \ABquestion{We have not introduced what resolved/unresolved 
  and so I think it is necessary to explain what they are exactly here or in the NLL derivation
  in Sec.\ref{sec:NLL-resum}?}

  \item Parton distribution functions enter through the ratios
  $f_i(x_\ell,v^{\frac{1}{a+b_\ell}}Q)/f_i(x_\ell,\mu_F)$,
  where $f_i(x_\ell,\mu_F)$ is the parton distribution function
  corresponding to the flavour of the parton $p_\ell$ entering the hard scattering, with
  momentum fraction $x_\ell$. The physical meaning of this ratio is the contribution 
  of all unresolved hard collinear emissions below the veto scale 
  $v^{\frac{1}{a+b_\ell}}Q$ gives the DGLAP evolution of a 
  parton distribution function. \ABquestion{I want
  to be a bit more explicit here, can I say:
  `` between the factorisation scale $\mu_F$ and the veto scale
  $v^{1/(a+b_\ell)}Q$, the incoming parton can emit collinear
  gluons that are too soft to be resolved by the observable.
  These unresolved emissions are not included in
  $\mathcal{F}_{\rm NLL}$ but they modify the effective PDF seen
  by the hard process. The ratio $f_i(x_\ell,v^{1/(a+b_\ell)}Q)/
  f_i(x_\ell,\mu_F)$ resums precisely this effect via DGLAP
  evolution from $\mu_F$ down to the veto scale.''}
  This contribution is absent in
  $e^+e^-$ annihilation, where there are no incoming partons, and
  is a genuinely new ingredient in hadronic collisions.
  \ABquestion{Is it necessary to derive how this comes about?}

\item Resolved soft and collinear emissions widely separated in angle,
  together with the appropriate virtual corrections, build up the
  NLL function $\mathcal{F}_{\rm NLL}(\lambda)$, derived in
  Sec.~\ref{sec:NLL-resum} and given in Eq.~\eqref{eq:FNLL-formula}.
  This ingredient is unchanged from the $e^+e^-$ case.

\item Finite virtual corrections of relative order $\alpha_s$ are
  encoded in the hard coefficient $H_{ij}^{(1)}$. This is a term of order
  $\alpha_s$ multiplying the NLL function, $\mathcal{F}^{(ij)}_{\rm NLL}(\lambda)$, 
  hence NNLL. For a generic process, $H_{ij}^{(1)}$ depends on $Q$, kinematic invariants and $\mu_R$. 
  \ndp{This contribution is process dependent and so will be entirely different between the $\ee$ and hadronic case.} 
  The general formula for $H_{ij}^{(1)}$ is derived in Sec.~\ref{sec:H1}; 
   
\item A single unresolved hard collinear splitting of leg
  $\ell$ combines with the corresponding virtual corrections and, in
  the case of initial-state radiation, to the collinear counterterm
  needed to renormalise the parton distribution functions. This gives
  the coefficient $C_{\mathrm{hc},\ell}^{(1)}$, which gives rise to a NNLL
  contribution since it is a term of order $\alpha_s$
  multiplying the NLL function, $\mathcal{F}^{(ij)}_{\rm NLL}(\lambda)$.
   \ABquestion{ I want to be explicit about this
  PDF counterterm as I have not explained it anywhere yet, 
  can I say something like ``The PDF counterterm is necessary to avoid double-counting.
  In the $\overline{\rm MS}$ scheme the PDF absorbs the collinear $1/\varepsilon$ divergences.
  Therefore, the real splitting in the collinear limit must be subtracted to prevent this divergence
  from being counted twice.  Once the real emission, virtual correction,
and PDF counterterm are combined, all $1/\varepsilon$ poles cancel
and one is left with a finite $C_{\mathrm{hc},\ell}^{(1)}$.''}
The final-state contribution is identical to that derived in Ref.~\cite{Arpino:2019ozn}, 
while the initial-state contribution is a new result of this work, derived
 in Sec.~\ref{sec:NNLL-initial}.  
  
  \item All remaining resolved emissions and their associated virtual
  corrections are collected in the NNLL function
  $\delta\mathcal{F}_{\rm NNLL}(\lambda)$. These consist of one
  ``special'' soft and/or collinear emission accompanied by an
  arbitrary ensemble of soft-collinear emissions widely
  separated in angle, together with the corresponding virtual
  corrections:
   \begin{equation}
    \label{eq:dFNNLL}
    \delta\mathcal{F}_{\rm NNLL} =
    \delta\mathcal{F}_{\rm sc}
    +\delta\mathcal{F}_{\rm hc}
    +\delta\mathcal{F}_{\rm rec}
    +\Delta\mathcal{F}_{\rm rec}
    +\delta\mathcal{F}_{\rm wa}
    +\Delta\mathcal{F}_{\rm wa}
    +\delta\mathcal{F}_{\rm correl}\,.
  \end{equation}
  Each term reflects the region of phase space spanned by the special emission
  and can be expressed as an integral over the soft-collinear measure $\dZ$ with the appropriate weight function $G(\{k_i\})$ 
  (see Eq.~\eqref{eq:dZ-def}). \ABquestion{Do I need to give a a more detailed
  explanation of what each correction does?}  
  
  The contributions of soft origin ---
$\delta\mathcal{F}_{\rm sc}$, $\delta\mathcal{F}_{\rm wa}$,
$\Delta\mathcal{F}_{\rm wa}$, and $\delta\mathcal{F}_{\rm correl}$
--- are identical to those in Ref.~\cite{Arpino:2019ozn} and
contain no new ISR contributions. This follows directly from the
matrix element factorisation in the soft limit established in
Sec.~\ref{subsec:soft-fact}. This describes how 
each soft emission matrix element depends only on the 
colour charge and direction of the emitting leg, 
not on its momentum fraction, flavour, or whether it is an
initial- or final-state parton. The soft corrections are
therefore universal between ISR and FSR. 


 The ISR contributions from the collinear region do change between ISR and FSR.
  The emissions that are collinear to an incoming leg carry a significant fraction of its
  momentum. They generate collinear divergences regulated by the PDF counterterm,
  and involve splitting functions convoluted with PDFs. Furthermore, they know
  about the identity of the parton and the momentum fraction it holds. This is again
  shown in Sec.~\ref{subsec:collinear-fact}, and is reflected in the hard-collinear contributions ---
  $\delta\mathcal{F}_{\rm hc}$, $\delta\mathcal{F}_{\rm rec}$, and
  $\Delta\mathcal{F}_{\rm rec}$.  These have the same formal expressions
  as in the $e^+e^-$ case, but their ingredients change for
  initial-state radiation.  In particular, for initial-state radiation, 
  these functions involve convolutions of splitting functions with PDFs.  
  These new ingredients
  are derived in Sec.~\ref{sec:NNLL-initial} and evaluated
  explicitly for one-jettiness in Sec.~\ref{sec:one-jettiness}.

  \end{itemize} 
  
  In the following two subsections we discuss $H_{ij}^{(1)}(Q,\mu_R)$
and $R_{\rm NNLL}(v)$ in more detail. The radiator can in principle
be taken directly from Ref.~\cite{Arpino:2019ozn}, since the colour
structure of $pp\to Z+\text{jet}$ is the same as the $e^+e^-$
three-jet case. Nevertheless we present its derivation here to
illustrate how the NLL radiator is extended to NNLL radiator and to
establish the conventions used throughout this chapter. The hard
function $H_{ij}^{(1)}$, by contrast, is always process-specific
and cannot be taken from Ref.~\cite{Arpino:2019ozn}. Its general
formula is derived in Sec.~\ref{sec:H1}.


\subsection{The hard coefficient $\mathbf{H_{ij}^{(1)}(Q,\mu_R)}$}
\label{sec:H1}

\ABquestion{What are the arguments $H_{ij}^{(1)}$ -- they keep changing throughout the text? } \\
\ABquestion{Do we need to mention about $\alpha_s^{\rm phys}$ here
(also in the radiator)? Also, are all of the $\alpha_s$'s functions
of $Q$ or $\mu_R$ here?}

\ndp{The hard coefficient $H_{ij}^{(1)}(Q)$ encodes the finite one-loop virtual
corrections to the hard scattering amplitude.} Its derivation sets the
conventions used throughout this chapter and introduces the web
function formalism that also underlies the radiator. We follow the
notation of Ref.~\cite{Catani:1998bh}, slightly adapted to our needs.

\ndp{At NNLL, we need the full one-loop virtual amplitudes. These
contain both UV divergences, which are regulated after renormalisation, and IR
divergences. Taking the UV renormalised virtual corrections, leaves us just 
with the IR divergences.}

\ABquestion{Do we need the full one-loop virtual corrections for NNLL?
In general what bits do we need for NNLL from the fixed-order calculation i.e.\
one loop, one real emission (NLO) pieces?}

\ABquestion{In resummation do we keep
all virtual corrections while the real emissions are restricted?}

The starting point to compute $H^{(1)}_{ij}$ is the following representation of UV-renormalised virtual corrections~\cite{Banfi:2018mcq,Arpino:2019ozn}:
\begin{multline}
  \label{eq:HQ}
  \mathcal{V}(Q;\varepsilon) = H_{ij}(\alpha_s(\mu_R),Q) \exp\left\{\int^{Q^2}\!\!\frac{dk^2}{k^2}\sum_{\ell=1}^3 \gamma_\ell\left(\alpha_s(k;\varepsilon)\right)\right\}\times \\ \times \exp\left\{-\int \frac{d^dk}{(2\pi)^d}\sum_{\ell=1}^3 \sum_{\bar \ell\ne \ell} w_{\ell \bar \ell}(m^2,m^2+\kappa_{\ell\bar \ell}^2;\varepsilon) \Theta\left(\frac 12\ln\frac{Q^2_{\ell\bar \ell}}{m^2+\kappa_{\ell\bar \ell}^2}-|\eta^{(\ell\bar \ell)}| \right) \Theta(Q_{\ell \bar \ell}-\kappa_{\ell\bar \ell})\right\}\,,
\end{multline}
where $Q_{\ell\bar\ell}^2=2p_\ell\cdot p_{\bar\ell}$ and dimensional
regularisation is used to regulate soft and collinear divergences,
with $d=4-2\varepsilon$.  
\ndp{The virtual corrections are split into hard-collinear virtual corrections which are governed
by the $\gamma_{\ell}$ exponential. The exponential with the web $w_{\ell\bar\ell}$  generates
soft virtual corrections where the web is the eikonal weight for a soft gluon exchanged between 
legs $\ell$ and $\bar\ell$, and the $\Theta$-functions restrict the integration to the region
where this gluon is genuinely soft (rapidity bounded by the dipole
kinematics and $\kappa_{\ell\bar\ell}<Q_{\ell\bar\ell}$). The hard
function $H_{ij}(\alpha_s,Q)$ then captures everything that is left
after these IR-divergent pieces are extracted -- it is the finite
remainder.}
The key ingredients are:
\begin{itemize}
\item
  the ``hard'' function $H_{ij}(\alpha_s,Q)$ for incoming channel $ij$, which has the following expansion in powers of $\alpha_s(Q)$:
  \begin{equation}
    \label{eq:HQ-exp}
    H_{ij}(\alpha_s) = 1+\frac{\alpha_s}{2\pi}H_{ij}^{(1)}(Q) +\left(\frac{\alpha_s}{2\pi}\right)^2H_{ij}^{(2)}(Q)+\dots 
  \end{equation}
  
\item the $d$-dimensional strong coupling $\alpha_s(k^2;\varepsilon)$, defined as the solution of the renormalisation-group equation
\begin{equation}
  \label{eq:RG-coupling}
  k^2\frac{d\alpha_s}{dk^2} = - \varepsilon \alpha_s+\beta(\alpha_s)\,,
\end{equation}
where $\beta(\alpha_s)$ is the QCD beta function, and renormalisation
is performed in the $\overline{\rm MS}$ scheme. The collinear anomalous dimension $\gamma_\ell(\alpha_s)$ has the expansion 
\begin{equation}
  \label{eq:gammal}
  \gamma_\ell (\alpha_s) = \gamma^{(0)}_\ell \frac{\alpha_s}{2\pi} + \gamma^{(1)}_\ell \left(\frac{\alpha_s}{2\pi} \right)^2+\dots
\end{equation}
where $\gamma_\ell^{(0)}$ is given in Eq.~\eqref{eq:gammal-flav}, and 
 \begin{equation}
  \label{eq:gammal1-flav}
  \gamma_\ell^{(1)} = \left\{
    \begin{split}
      &\frac{C_F}{2}\left(C_F\left(\frac{3}{4}-\pi^2+12\zeta_3\right)+C_A\left(\frac{17}{12}+\frac{11\pi^2}{9}-6\zeta_3\right)-n_f\left(\frac{1}{6}+\frac{2\pi^2}{9}\right)\right)\,,\quad \text{if $\ell$ is a quark,} \\
      & C_A^2\left(\frac{8}{3}+3\zeta_3\right)-\frac{n_f}{2}C_F-\frac{2}{3}n_fC_A \,,\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\text{if $\ell$ is a gluon}\,;
    \end{split}
    \right.
\end{equation}

\item the ``web'' function
  $w_{\ell \bar \ell}(m^2,m^2+\kappa_{\ell\bar \ell}^2;\varepsilon)$
  which consistently resum the contribution of multiple soft real
  emissions from the dipole $(\ell\bar\ell)$. Physically, the web function is the weight associated
  with the emission of a gluon of mass $m$, inclusive with respect to
  all its splittings. Since Feynman rules are invariant under
  rescaling of the emitters' four-momenta, the web function can only
  depend on variables that are invariant under such rescalings. These
  are $m^2$ and $\kappa_{\ell\bar\ell}^2$ (defined in Eq.~\eqref{eq:kappa}).
  The web function admits an expansion in
  powers of the strong coupling,
  $w_{\ell \bar \ell} = \alpha_s(\mu_R) w_{\ell \bar
    \ell}^{(1)}+\alpha_s^2(\mu_R) w_{\ell \bar \ell}^{(2)}+\dots$, and
  \begin{equation}
    \label{eq:w1}
    w_{\ell \bar \ell}^{(1)}(m^2,m^2+\kappa^2;\varepsilon) = (4\pi)^2\frac{C_{\ell\bar\ell}}{\kappa_{\ell\bar\ell}^2}\,\mu_R^{2\epsilon}\frac{e^{\varepsilon\gamma_E}}{(4\pi)^{\varepsilon}}\,\delta(m^2)\,,
  \end{equation}
  where $C_{\ell\bar \ell}$ is the colour dipole factor introduced in Sec.~\ref{subsubsec:Rs}. \ndp{The $1/\kappa_{\ell\bar\ell}^2$ is the eikonal factor for soft gluon emission from the dipole with colour factor $C_{\ell\bar\ell}$ which occurs in soft matrix element factorisation as discussed in Sec.~\ref{subsec:soft-fact}. The term $\delta(m^2)$ is what places the gluon on-shell};
\item the $d$-dimensional phase space, given by
  \begin{equation}
    \label{eq:dk}
    \frac{d^d k}{(2\pi)^d} = \frac{d\eta^{(\ell\bar \ell)}}{2\pi}\, dm^2 \,\frac{d\kappa^2_{\ell\bar \ell}}{(4\pi)^2}\left(\frac{\pi}{\kappa^2_{\ell\bar\ell}}\right)^\varepsilon\frac{\Gamma(1-\varepsilon)}{\Gamma(1-2\varepsilon)}\frac{d\phi}{\pi}\,(\sin^2\phi)^{-\varepsilon}\,\Theta(\sin\phi)\,. 
  \end{equation}
\end{itemize}

We extract $H_{ij}^{(1)}$ by expanding both sides of
Eq.~\eqref{eq:HQ} to order $\alpha_s$ and comparing.
Evaluating the web function integral gives
\begin{multline}
  \label{eq:w-exp}
  -\alpha_s(\mu_R)  \int \frac{d^dk}{(2\pi)^d} w^{(1)}_{\ell \bar \ell}(m^2,m^2+\kappa_{\ell\bar \ell}^2;\varepsilon) \Theta\left(\frac 12\ln\frac{Q^2_{\ell\bar \ell}}{m^2+\kappa_{\ell\bar \ell}^2}-|\eta^{(\ell\bar \ell)}| \right) \Theta(Q_{\ell \bar\ell}-\kappa_{\ell\bar \ell}) \\ = -\frac{\alpha_s(\mu_R)}{2\pi}
  \left(\frac{\mu_R^2}{Q^2_{\ell\bar \ell}}\right)^\varepsilon \frac{e^{\varepsilon\gamma_E}}{\Gamma(1-\varepsilon)} \frac{C_{\ell\bar \ell}}{\varepsilon^2}\,,
\end{multline}
and the anomalous dimension integral gives
\begin{equation}
  \label{eq:gamma0-int}
  \int_0^{Q^2}\!\!\frac{dk^2}{k^2}\gamma_\ell^{(0)}\frac{\alpha_s(k^2;\varepsilon)}{2\pi} = \frac{\alpha_s(\mu_R)}{2\pi}\gamma_\ell^{(0)} \int_0^{Q^2}\!\!\frac{dk^2}{k^2}\left(\frac{\mu_R^2}{k^2}\right)^{\varepsilon} = -\frac{1}{\varepsilon}\left(\frac{\mu_R^2}{Q^2}\right)^{\varepsilon} \gamma_\ell^{(0)}\frac{\alpha_s(\mu_R)}{2\pi}\,.
\end{equation}
\ndp{The poles in $\varepsilon$ from the two exponentials are precisely
the IR poles of the full amplitude predicted by the Catani
formula~\cite{Catani:1998bh}, and $H_{ij}^{(1)}$ is the associated
finite part.}
Using the fact that $\mathcal{V}(Q;\varepsilon)$ has the universal structure~\cite{Catani:1998bh}\footnote{Note that the general formula of~\cite{Catani:1998bh} includes also phase factors (Coulomb phases) that arise when the two legs $\ell$ and $\bar \ell$ are both incoming or outgoing. In the case of three emitters, these phase factors give rise to a finite contribution, included in $\mathcal{H}_1$.}
\begin{equation}
  \label{eq:HQ-expanded}
  \mathcal{V}(Q;\varepsilon) = 1+\frac{\alpha_s(\mu_R)}{2\pi}\left[-\frac{e^{\varepsilon\gamma_E}}{\Gamma(1-\varepsilon)}\sum_\ell\left(\frac{1}{\varepsilon^2}+\frac{1}{\varepsilon}\frac{\gamma_\ell^{(0)}}{C_\ell}\right)\sum_{\bar \ell\ne \ell} C_{\ell\bar\ell}\left(\frac{\mu_R^2}{Q_{\ell\bar \ell}^2}\right)^{\varepsilon}+\mathcal{H}_1\right]\,,
\end{equation}
and comparing the above result with the right-hand side of Eq.~(\ref{eq:HQ}),
\begin{equation}
  \label{eq:HQ-rep-expanded}
  \mathcal{V}(Q;\varepsilon) = 1+\frac{\alpha_s(\mu_R)}{2\pi}\left\{-\frac{e^{\varepsilon\gamma_E}}{\Gamma(1-\varepsilon)}\sum_\ell\left[\sum_{\bar \ell\ne \ell}\left(\frac{\mu_R^2}{Q^2_{\ell\bar\ell}}\right)^{\varepsilon}\frac{C_{\ell\bar \ell}}{\varepsilon^2}+\left(\frac{\mu_R^2}{Q^2}\right)^{\varepsilon}\frac{\gamma_\ell^{(0)}}{\varepsilon}\right]+H_1\right\}\,,
\end{equation}
we find
\begin{equation}
  \label{eq:H1-final}
  H_{ij}^{(1)}(Q) = \mathcal{H}_1+ \sum_\ell\frac{\gamma_\ell^{(0)}}{C_\ell}\sum_{\bar \ell\ne \ell} C_{\ell\bar \ell}\ln\left(\frac{Q^2_{\ell\bar\ell}}{Q^2}\right)
    \,.
\end{equation}
\ndp{The result splits naturally into two pieces. The logarithmic terms
$\sum_\ell(\gamma_\ell^{(0)}/C_\ell)\sum_{\bar\ell}C_{\ell\bar\ell}
\ln(Q^2_{\ell\bar\ell}/Q^2)$ are universal --- they depend only on
the colour charges and kinematics of the hard legs, not on the
specific process. The finite constant $\mathcal{H}_1$ is
process-specific, where it is determined by the one-loop amplitude for
the hard subprocess and must be computed separately for each
process.}


\subsection{The NNLL Radiator}
\label{sec:radiator}

The NNLL radiator extends the NLL result of
Sec.~\ref{subsec:NLL-radiator}, and contains both NLL and NNLL contributions. Here we retain the same notation and kinematic
variables used throughout. The radiator is obtained
operationally from the singular part of the virtual corrections in
Eq.~(\ref{eq:HQ}), by cutting off hard-collinear contributions down to
the scale $Qv^{1/(a+b_\ell)}$, and including suitably defined
``unresolved'' soft emission. \ndp{The web
function integral of Eq.~\eqref{eq:w-exp} and the anomalous
dimension integral of Eq.~\eqref{eq:gamma0-int}, restricted to this
soft region, give rise to the soft and hard-collinear pieces
respectively.} This gives rise to three new contributions that appear at 
NNLL that were absent at NLL:
\begin{itemize}
\item the mass correction $\delta
R_{\ell\bar\ell}$ arising from the finite gluon mass in the web
function
\item the second-derivative term $R''$ in the azimuthal
expansion \add{Add in Eq.}
\item the two-loop anomalous dimension $\gamma_\ell^{(1)}$
in the hard-collinear piece.
\end{itemize} 

The NNLL splits into a soft and hard-collinear contribution as in NLL,
\begin{equation}
  \label{eq:radiator-NNLL}
  R_{\rm NNLL}(v) = R_{\rm s}(v)+ R_{\rm hc}(v)\,,
  \end{equation}
neglecting all terms beyond NNLL accuracy.
  
  The soft radiator is
  \begin{equation}
  	R_{\rm s}(v)= \sum_\ell\sum_{\bar \ell\ne \ell} C_{\ell\bar \ell} \,\mathcal{R}_{\ell\bar\ell}(v)\,, 
  \end{equation}
  where $\mathcal{R}_{\ell\bar\ell}(v)$ is obtained from the web
function integral restricted to unresolved soft emissions, giving rise to the finite contributions
\begin{multline}
  \label{eq:Rll}
  \mathcal{R}_{\ell\bar\ell}(v) = \int\frac{d^4
    k}{(2\pi)^4}w_{\ell\bar\ell}(m^2,m^2+\kappa_{\ell\bar\ell}^2;0)
  \Theta\left(\frac 12\ln\frac{Q^2_{\ell\bar
        \ell}}{m^2+\kappa_{\ell\bar \ell}^2}-\eta^{(\ell\bar \ell)}
  \right) \Theta(Q_{\ell \bar\ell}-\kappa_{\ell\bar \ell}) \times \\
  \times \Theta(\eta^{(\ell\bar \ell)})
  \Theta\left(d^{(\ell\bar\ell)}_\ell\left(\frac{\kappa_{\ell\bar
          \ell}}{Q_{\ell\bar\ell}}\right)^{a}
    e^{-b_\ell\eta^{(\ell\bar \ell)}} g_\ell(\phi)-v\right)\,.
\end{multline}
\ndp{The key difference from the NLL soft radiator of
Eq.~\eqref{eq:Rll-dip} is that the web function $w_{\ell\bar\ell}$
here retains its full $m^2$-dependence rather than being set to
$m^2=0$.}

We note that, in the above expression, except the physical bound on
rapidity
$\eta^{(\ell\bar \ell)}<\ln(Q_{\ell\bar
  \ell}/\sqrt{m^2+\kappa_{\ell\bar \ell}^2})$, all other constraints
are conventional. Changing them will affect the expression of $\FNLL$
and $\delta\FNNLL$, leading to an expression of $\Sigma_{ij}(v)$ that
differs from the one obtained using Eq.~(\ref{eq:Rll}) by N$^3$LL
corrections. 

We can account for the $\phi$ and $m^2$ dependence of the observable at NNLL accuracy, by reorganising $\mathcal{R}_{\ell\bar\ell}(v)$ as follows
\begin{equation}
\label{eq:Rll-rearranged}
  \mathcal{R}_{\ell\bar\ell}(v) = \int\frac{d\phi}{2\pi}\, R_{\ell\bar \ell}\left(\frac{v}{d_{\ell\bar \ell} \,g_\ell(\phi)},Q_{\ell\bar \ell}\right)+\delta R_{\ell\bar \ell}(v)\,,
\end{equation}
where at NNLL accuracy \ABquestion{Cannot quite make sense 
of the argument $v/d_{\ell\bar \ell}g_{\ell}$?}
\begin{multline}
  \label{eq:Rll-expanded}
\int\frac{d\phi}{2\pi}\, R_{\ell\bar \ell}\left(\frac{v}{d_{\ell\bar \ell} \,g_\ell(\phi)},Q_{\ell\bar \ell}\right)\simeq  
R_{\ell\bar \ell}\left(v,Q_{\ell\bar \ell}\right)\\+\langle \ln(d_{\ell\bar \ell}\,g_\ell)\rangle R'_{\ell\bar \ell}\left(v,Q_{\ell\bar \ell}\right)+\frac 12
\langle \ln^2(d_{\ell\bar \ell}\,g_\ell)\rangle R''_{\ell\bar \ell}\left(v,Q_{\ell\bar \ell}\right)\,,
\end{multline}
and $R_{\ell\bar \ell}(v,Q_{\ell\bar \ell})$ is exactly the same as in the NLL radiator in the previous chapter given by Eq.~\eqref{eq:Rll-final}, $\langle\cdots\rangle$ denotes the azimuthal average of
Eq.~\eqref{eq:azimuthal-average}, $R'_{\ell\bar\ell}\equiv
-vdR_{\ell\bar\ell}/dv$, and $R''_{\ell\bar\ell}\equiv
-vdR'_{\ell\bar\ell}/dv$. 
The ``mass correction'' $\delta R_{\ell\bar \ell}(v)$ in Eq.~\eqref{eq:Rll-rearranged}
is given by
\begin{multline}
  \label{eq:dRll}
  \delta R_{\ell\bar\ell}(v) = -\int\frac{d^4
    k}{(2\pi)^4}w_{\ell\bar\ell}(m^2,m^2+\kappa_{\ell\bar\ell}^2;0)
 \Theta(Q_{\ell \bar\ell}-\kappa_{\ell\bar \ell}) \Theta(\kappa_{\ell\bar\ell}-v^{\frac1{a+b_\ell}} Q_{\ell\bar \ell}) \times \\
  \times \Theta\left(\eta^{(\ell\bar \ell)}-\frac 12\ln\frac{Q^2_{\ell\bar
        \ell}}{m^2+\kappa_{\ell\bar \ell}^2}
  \right)\Theta\left(\ln\frac{Q_{\ell\bar
        \ell}}{\kappa_{\ell\bar \ell}}-\eta^{(\ell\bar \ell)}  \right)\,.
\end{multline}



\ndp{The hard-collinear radiator resums virtual corrections collinear to
each leg, integrated from the veto scale $Qv^{1/(a+b_\ell)}$ up to
$Q$, as at NLL (cf.\ Eq.~\eqref{eq:Rhc-NLL}), but now including
the two-loop anomalous dimension:} 
\begin{equation}
  \label{eq:Rhc-NNLL}
  R_{\rm hc}(v) =
  -\sum_\ell\int^{Q^2}_{Q^2v^{\frac{2}{a+b_\ell}}}
  \!\!\frac{dk^2}{k^2}\,\gamma_\ell\!\left(\alpha_s(k)\right)\,,
\end{equation}
where $\gamma_\ell[\alpha_s]$ is given in Eq.~\eqref{eq:gammal}. Retaining only
$\gamma_\ell^{(0)}$ recovers the NLL result of
Eq.~\eqref{eq:Rhc-NLL}, while the $\gamma_\ell^{(1)}$ term is the new
NNLL contribution. 

Just as in the case for the NLL radiator, we use the ansatz
in Eq.~\eqref{eq:ansatz}, and find the expressions below:
\begin{equation}
	R_{\ell\bar\ell}(v)=-\frac{\lambda}{\alpha_s(Q)\beta_0}g_1^{(\ell)}(\lambda)-g_2^{(\ell)}(\lambda)-\mathfrak{g}_2^{(\ell)}(\lambda,Q_{\ell\bar\ell})-\frac{\alpha_s(Q)}{\pi}\left(g_3^{(\ell)}(\lambda)+\mathfrak{g}_3^{(\ell)}(\lambda,Q_{\ell\bar\ell})\right)\,,
\end{equation}

\begin{equation}
	\delta R_{\ell\bar\ell}(v)=-\frac{\alpha_s(Q)}{\pi}\delta g_3^{(\ell)}(\lambda)\,,
\end{equation}

\begin{equation}
	R_{\mathrm{hc},\ell}(v)=-h_2^{(\ell)}(\lambda)-\frac{\alpha_s(Q)}{\pi}h_3^{(\ell)}(\lambda)\,,
\end{equation}
where $g_1^{(\ell)}(\lambda), g_2^{(\ell)}(\lambda), \mathfrak{g}_2^{(\ell)}(\lambda,Q_{\ell\bar\ell}),$ and $h_2^{(\ell)}(\lambda)$ can be found in Eq.~\eqref{eq:RNLL-g-h} and the remaining terms are expressed below:


\begin{subequations}
\begin{align}
	&g_3^{(\ell)}(\lambda)=\left[K^{(1)}\frac{\beta_1\left(a^2(a+b_{\ell}-2\lambda)\ln\left(1-\frac{2\lambda}{a}\right)-(a+b_{\ell})^2(a-2\lambda)\ln\left(1-\frac{2\lambda}{a+b_{\ell}}\right)+6b_{\ell}\lambda^2\right)}{4\pi b_{\ell}\beta_0^3(a-2\lambda)(a+b_{\ell}-2\lambda)}\right.\notag \\
	&\qquad\qquad+\frac{\left(\beta_1^2(a+b_{\ell})^2(a-2\lambda)\ln^2\left(1-\frac{2\lambda}{a+b_{\ell}}\right)-4b_{\ell}\lambda^2\left(\beta_0\beta_2+\beta_1^2\right)\right)}{4\pi b_{\ell}\beta_0^4(a-2\lambda)(a+b_{\ell}-2\lambda)} \notag \\
	&\qquad\qquad-\frac{a\ln\left(1-\frac{2\lambda}{a}\right)\left(2\beta_0\beta_2(a-2\lambda)+a\beta_1^2\ln\left(1-\frac{2\lambda}{a}\right)+4\beta_1^2\lambda\right)}{4b_{\ell}\beta_0^4(a-2\lambda)} \notag\\
	&\qquad\qquad+\frac{(a+b_{\ell})\ln\left(1-\frac{2\lambda}{a+b_{\ell}}\right)\left(\beta_0\beta_2(a+b_{\ell}-2\lambda)+2\beta_1^2\lambda\right)}{2b_{\ell}\beta_0^4(a+b_{\ell}-2\lambda)} \notag\\
	&\left.\qquad\qquad -K^{(2)}\frac{2\lambda^2}{8\pi^2(a-2\lambda)(a+b_{\ell}-2\lambda)\beta_0^2} \right]\,, \\[6pt]
&\mathfrak{g}_3^{(\ell)}(\lambda,Q_{\ell\bar\ell})=\pi\beta_0\lambda^2\left(\frac{dg_1^{(\ell)}(\lambda)}{d\lambda}+\frac{\lambda}{2}\frac{d^2g_1^{(\ell)}(\lambda)}{d\lambda^2}\right)\ln^2\left(\frac{Q_{\ell\bar\ell}^2}{Q^2}\right)\notag \\[6pt]
&\qquad\qquad -\pi\beta_0\lambda\left(\frac{dg_2^{(\ell)}(\lambda)}{d\lambda}+\frac{\beta_1}{\beta_0^2}\lambda\frac{dg_1^{(\ell)}(\lambda)}{d\lambda}\right)\ln\left(\frac{Q_{\ell\bar\ell}^2}{Q^2}\right)\,, \\[6pt]
&\delta g_3^{(\ell)}(\lambda)=-\zeta_2\frac{\lambda}{(a+b_{\ell}-2\lambda)}\,,\\[6pt]
&h_3^{(\ell)}(\lambda)=-\gamma_{\ell}^{(0)}\frac{\beta_1\left((a+b_{\ell})\ln\left(1-\frac{2\lambda}{a+b_{\ell}}\right)+2\lambda\right)}{2\beta_0^2(a+b_{\ell}-2\lambda)}
+\gamma_{\ell}^{(1)}\frac{\lambda}{2\pi\beta_0(a+b_{\ell}-2\lambda)}\,,
\end{align}
\end{subequations}
where $\beta_0, \beta_1, \beta_2$ can be found in Eq.~\eqref{eq:beta}
and $\gamma_{\ell}^{(0)}, \gamma_{\ell}^{(1)}$ can be found in
Eqs.~\eqref{eq:gammal-flav} and.~\eqref{eq:gammal1-flav} 
respectively.

\todo{add the derivative of the radiators in here.}
We can express derivatives of the NNLL radiator as a sum over legs $\ell$ to be:
\begin{subequations}
\begin{align}
	&R'_{\rm NNLL}(\lambda)= \\
	&R''_{\rm NNLL}(\lambda)= \\
	&\Delta R'_{\ell, \mathrm{NNLL}}(\lambda)=\,.
	\end{align}
\end{subequations}

\section{NNLL resummation: initial-state radiation}
\label{sec:NNLL-initial}

\todo{Sort out that the momentum fractions are consistent.}
A general treatment of initial-state radiation at NNLL is currently
missing. Results exist for $N$-jettiness both in deep-inelastic
scattering and hadron collisions \add{References missing}, as well as
for the transverse momentum of a colour singlet or a top-antitop pair,
and for selected observables in colour-singlet
production~\cite{vanBeekveld:2025zjh}.

We restrict ourselves to event-shape variables and, as is customary in
the \textsc{ARES} approach, we identify the relevant emissions contributing 
at NNLL: a single hard and collinear emission $k$, accompanied by an arbitrary number of soft and collinear emissions $k_1,\dots,k_n$. 
The emission $k$ carries a fraction $(1-z)/z$ of the 
momentum $p_\ell$ entering the hard
scattering, 
\begin{equation}
  \label{eq:k||l}
  k = \frac{1-z}{z} p_\ell + k_t^{(\ell)}+ \mathcal{O}\left(\left(k_t^{(\ell)}\right)^2\right)\,.
\end{equation}
To extend \textsc{ARES} to initial-state radiation, we follow the 
same principles as for final-state radiation: 
\begin{itemize}
\item we consider one collinear emission $k$ and the corresponding virtual
  corrections, accompanied by an arbitrary
  number of soft and collinear emissions $\{k_i\}$;
\item we subtract the double counting occurring when $k$ is also soft,
\ndp{since such contributions are already included at NLL in $\FNLL$.}
\end{itemize}
The procedure will involve contributions that are not separately
finite in four dimensions. We use dimensional regularisation to make
each contribution finite in $4-2\varepsilon$ dimensions. For initial-state radiation, virtual corrections must include the
counterterm arising from the renormalisation of PDFs. 
\question{Why must we include this?}
Throughout, we assume that soft emissions coherently factorise 
from collinear radiation, i.e.\ a soft gluon sees only the total 
colour charge of harder emissions at smaller angles. 
Recently identified
coherence-violating corrections~\cite{Banfi:2025mra,Becher:2026kbr}
first enter at order $\alpha_s^4$ and require a dedicated treatment
beyond the scope of this thesis.

We need to distinguish two cases, according to whether the parton
entering the hard scattering is a quark or a gluon.

\subsection{Quark entering the hard scattering}
\label{sec:quark-in}

Two collinear splittings can produce a quark entering the
hard scattering: an incoming quark radiating a collinear
gluon ($q\to qg$), and an incoming gluon splitting into a
quark-antiquark pair ($g\to q\bar q$). 
For $Z$+jet production these are illustrated in
Figs.~\ref{fig:q-incoming-qqg} and~\ref{fig:q-incoming-gqq}.

\begin{figure}[H]
\centering

\begin{subfigure}{0.48\textwidth}
\centering
\begin{tikzpicture}[scale=0.85]
\begin{feynman}
  \vertex (H) at (0,0);
  \vertex (v1) at (-2.0, 1.5);
  \vertex (q0) at (-4.2, 3.0);
  \vertex (qbar) at (-4.2, -2.2);
  \vertex (gISR) at (0.6, 2.95);
  \vertex (Z)   at (3.8,  1.6);
  \vertex (gout) at (3.8, -1.6);
  \diagram*{
    (q0)   -- [fermion] (v1),
    (v1)   -- [fermion, edge label={$z$}] (H),
    (v1)   -- [gluon]                     (gISR),
    (qbar) -- [anti fermion]              (H),
    (H)    -- [boson]                     (Z),
    (H)    -- [gluon]                     (gout),
  };
  \draw[fill=white] (H) circle (0.8cm) node {$\mathcal{H}_{q\bar{q}}$};
\end{feynman}
\node[above left]  at (-4.2, 3.0)  {$q$};
\node[above right] at (-2.1, 2.3)  {$1-z$};
\node[below left]  at (-4.2, -2.2) {$\bar{q}$};
\node[above right] at (3.8,  1.6)  {$Z$};
\node[below right] at (3.8, -1.6)  {$g$};
\end{tikzpicture}
\caption{$q \to qg$ splitting}
\label{fig:q-incoming-qqg}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\begin{tikzpicture}[scale=0.85]
\begin{feynman}
  \vertex (H) at (0,0);
  \vertex (v1) at (-2.0, 1.5);
  \vertex (q0) at (-4.2, 3.0);
  \vertex (qbar) at (-4.2, -2.2);
  \vertex (gISR) at (0.6, 2.95);
  \vertex (Z)   at (3.8,  1.6);
  \vertex (gout) at (3.8, -1.6);
  \diagram*{
    (q0)   -- [gluon] (v1),
    (v1)   -- [fermion, edge label={$z$}] (H),
    (v1)   -- [anti fermion]              (gISR),
    (qbar) -- [anti fermion]              (H),
    (H)    -- [boson]                     (Z),
    (H)    -- [gluon]                     (gout),
  };
  \draw[fill=white] (H) circle (0.8cm) node {$\mathcal{H}_{q\bar{q}}$};
\end{feynman}
\node[above left]  at (-4.2, 3.0)  {$g$};
\node[above right] at (-2.1, 2.3)  {$1-z$};
\node[below left]  at (-4.2, -2.2) {$\bar{q}$};
\node[above right] at (3.8,  1.6)  {$Z$};
\node[below right] at (3.8, -1.6)  {$g$};
\end{tikzpicture}
\caption{$g \to q\bar{q}$ splitting}
\label{fig:q-incoming-gqq}
\end{subfigure}

\caption{Initial-state radiation contributions to the $q\bar{q}$ channel of $Z$+jet production, from the $q \to qg$ splitting (left)and the $g \to q\bar{q}$ splitting (right).  The hard scattering is denoted by $\mathcal{H}_{q\bar{q}}$, where the subscripts indicate the incoming partons.}
\label{fig:q-incoming}
\end{figure}

We first discuss the $q\to qg$ splitting in detail and then generalise to $g\to q\bar q$. We consider an incoming quark carrying a fraction $x/z$ of the proton momentum, which emits a collinear gluon carrying a fraction $1-z$ of its momentum. After the emission, the quark enters the hard scattering
with a fraction $x$ of the proton momentum. The contribution of this emission to $\Sigma_{ij}(\{\tilde p\},v)$ defines a function $\mathcal{F}_{qq}$, which is reorganised to give various NNLL contributions. Our starting point is
\begin{equation}
  \label{eq:FISR-start}
  \begin{split}
    \mathcal{F}_{qq} & = \int \dZ \left\{\int_0^{Q^2}\frac{dk_t^2}{k_t^2}\left(\frac{\mu_R^2}{k_t^2}\right)^{\varepsilon} \frac{\alpha_s(k_t)}{2\pi}\int_0^{2\pi}\frac{d\phi}{2\pi}\times\right. \\& \times \left[\int_{x_\ell}^1 \frac{dz}{z}\hat P_{qq}(z;\varepsilon)\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\}}{v}\right) \frac{f_q(x_\ell/z,\mu_F)}{f_q(x_\ell,\mu_F)}- \int_0^1 dz \frac{2C_\ell}{1-z}\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\}}{v}\right)\right] \\
   & \left[-\gamma_\ell^{(0)}\int_0^{Q^2v^{2/(a+b_\ell)}}\frac{dk^2}{k^2}\left(\frac{\mu_R^2}{k^2}\right)^{\varepsilon}\frac{\alpha_s(k)}{2\pi}
     +\frac{1}{\varepsilon} \left(\frac{\mu_R^2}{\mu_F^2}\right)^{\varepsilon} \frac{\alpha_s(\mu_R)}{2\pi} \int_{x_\ell}^1 \frac{dz}{z} P_{qq}(z) \frac{f_q(x_\ell/z,\mu_F)}{f_q(x_\ell,\mu_F)}\right] \times \\ & \times \left. \Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\}}{v}\right)\right\} \,.
\end{split}
\end{equation}
where we have introduced the regularised Altarelli-Parisi (AP) splitting function
\begin{equation}
  \label{eq:regularised-Pij}
  \hat P_{ij}(z;\varepsilon) = \hat P_{ij}(z;0)+\varepsilon P^{(\varepsilon)}_{ij}(z)\,, \qquad i=q,g\,. 
\end{equation}
In the present case,
\begin{equation}
  \label{eq:Pqq}
  \hat P_{qq}(z;0)  = C_F \frac{1+z^2}{1-z}\,, \qquad \hat P_{qq}^{(\varepsilon)}(z)=-C_F (1-z)\,,
\end{equation}
and $P_{qq}(z) = (\hat{P}_{qq}(z;0))_+$ is the full AP splitting function.

We now comment on the physical meaning of the ingredients within Eq.~\eqref{eq:FISR-start}.
The second line represents the splitting of a
quark emitting a collinear gluon $k$ in $4-2\varepsilon$ dimensions,
minus the contribution when the gluon is soft, already accounted for
in $\mathcal{F}_{\rm NLL}$. The third line contains virtual
corrections, as well as the collinear counterterm needed to renormalise
the parton distribution function. The parameter $b_\ell$ corresponds
to the leg identified by the quark entering the hard process.  The
expression in Eq.~\eqref{eq:FISR-start} contains single logarithms of
the form $\alpha_s\ln[Qv^{1/(a+b_\ell)}/\mu_R]$ and
$\alpha_s\ln[Qv^{1/(a+b_\ell)}/\mu_F]$. These logarithms can be
resummed by fixing $\mu_R=\mu_F=Qv^{1/(a+b_\ell)}$. \ABquestion{Are these
resummed or do they simply just disappear?} \ndp{At NNLL, we can set
$\alpha_s(k)=\alpha_s(k_t)=\alpha_s(Qv^{1/(a+b_\ell)})$ and obtain a
finite expression, up to subleading terms of order $\alpha_s^2$.}


In order to compute $\mathcal{F}_{qq}$ in four dimensions, we add and subtract
a counterterm, $\mathrm{CT}$, which has the following form:
\begin{multline}
    \mathrm{CT}=\int \dZ \frac{\alpha_s(Qv^{1/(a+b_\ell)})}{2\pi} \int\frac{dk_t^2}{k_t^2}\left(\frac{\mu_R^2}{k_t^2}\right)^{\varepsilon} \int_0^{2\pi}\frac{d\phi}{2\pi} \times \\
    \times \left[\int_{x_\ell}^1\frac{dz}{z}\hat P_{qq}(z;\varepsilon)\frac{f_q(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_q(x_\ell,Qv^{\frac{1}{a+b_\ell}})}-\int_0^1dz\frac{2 C_\ell}{1-z}\right]\times \\ \times \Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k)}{v}\right)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\}}{v}\right)\,.
  \end{multline}
  Adding and subtracting $\mathrm{CT}$ recasts $\mathcal{F}_{qq}$ in
the form of Eq.~\eqref{eq:Sigma-NNLL}:
\begin{equation}
  \label{eq:FISR-split}
  \mathcal{F}_{qq} = \frac{\alpha_s(Q)}{\pi}\left( \delta \mathcal{F}_{\rm rec,qq}+\delta\mathcal{F}_{\rm hc,qq}\right)+ \frac{\alpha_s(Q v^{1/(a+b_\ell)})}{2\pi} C^{(1)}_{\rm hc,qq} \mathcal{F}_{\rm NLL}\,.
\end{equation}
The coefficient $C^{(1)}_{\rm hc,qq}$ plays the same role as an integrated counterterm in fixed-order calculations, whilst the sum $\delta \mathcal{F}_{\rm rec,qq}+\delta\mathcal{F}_{\rm hc,qq}$ represents the subtracted contribution of real radiation. \ABquestion{Don't understand what we mean by \emph{subtracted contribution of real radiation}?} We now proceed to the detailed derivation of Eq.~\eqref{eq:FISR-split}. 

The coefficient $C^{(1)}_{\mathrm{hc, qq}}$ is obtained by adding $\mathrm{CT}$ to the last two lines of Eq.~\eqref{eq:FISR-start}:
\begin{equation}
\begin{aligned}
  & C^{(1)}_{\rm hc,qq} \mathcal{F}_{\rm NLL}(\lambda)= \int \dZ \int_0^{2\pi}\frac{d\phi}{2\pi}\,\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right) \times \\ & \times \left\{ -\gamma_\ell^{(0)}\int_0^{Q^2v^{2/(a+b_\ell)}}\frac{dk^2}{k^2}\left(\frac{Q^2v^{2/(a+b_\ell)}}{k^2}\right)^{\varepsilon}+\frac{1}{\varepsilon} \int_{x_\ell}^1 \frac{dz}{z} P_{qq}(z) \frac{f_q(x_\ell/z,Q v^{1/(a+b_\ell)})}{f_q(x_\ell,Q v^{1/(a+b_\ell)})} \right. \\ 
   &\left. +\int_0^{Q^2}\frac{dk_t^2}{k_t^2}\left(\frac{Q^2v^{2/(a+b_\ell)}}{k_t^2}\right)^{\varepsilon} \left[\int_{x_\ell}^1\frac{dz}{z} \hat P_{qq}(z;\varepsilon)\frac{f_q(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_q(x_\ell,Qv^{\frac{1}{a+b_\ell}})}-\int_0^1dz\frac{2 C_\ell}{1-z}\right]\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k)}{v}\right)\right\}\,.
\end{aligned}
\end{equation}
By comparing the first term on the right-hand-side of the above
equation to Eq.~\eqref{eq:FNLL-formula}, we recognise $\FNLL(\lambda)$. Using the parametrisation of
$V_{\mathrm{sc}}(\{\tilde{p}\},k)$ in terms of the splitting variables
\ABquestion{How do we get this parameterisation?}
\begin{equation}
\label{eq:Vsc-param}
V_{\mathrm{sc}}(\{\tilde{p}\},k)=d_\ell g_\ell(\phi) \left(\frac{Q}{2 E_\ell}\right)^{b_\ell}\left(\frac{k_t}{Q}\right)^{a+b_\ell}\frac{1}{(1-z)^{b_\ell}}\,,
\end{equation}
together with $\gamma_\ell^{(0)}=-3/2 C_F$, we find 
\begin{multline}
C^{(1)}_{\rm hc, qq}=  \int_{x_\ell}^1\frac{dz}{z}\,\frac{f_q(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_q(x_\ell,Qv^{\frac{1}{a+b_\ell}})}\times \\ \times
C_F\left[\frac{2 }{a+b_\ell}\left(b_\ell\left(\frac{\ln(1-z)}{1-z}\right)_+-\frac{1}{(1-z)_+}
        \left(\langle\ln d_\ell g_\ell\rangle-b_\ell \ln\frac{2E_\ell}{Q}\right)\right)(1+z^2)+(1-z)\right]\,.
  \end{multline}
  Using the regularised AP function in Eq.~\eqref{eq:Pqq}, the above expression can be written in the more compact form
  \begin{multline}
    \label{eq:C1qq-final}
C^{(1)}_{\rm hc, qq}=  \int_{x_\ell}^1\frac{dz}{z}\,\frac{f_q(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_q(x_\ell,Qv^{\frac{1}{a+b_\ell}})}\times \\ \times
\left[\frac{2 }{a+b_\ell}\left(b_\ell\left(\frac{\ln(1-z)}{1-z}\right)_+-\frac{1}{(1-z)_+}
        \left(\langle\ln d_\ell g_\ell\rangle-b_\ell \ln\frac{2E_\ell}{Q}\right)\right)(1-z)\hat P_{qq}(z;0)-P_{qq}^{(\varepsilon)}(z)\right]\,.
  \end{multline}
  Referring back to Eq.~\eqref{eq:FISR-split}, we now evaluate the remaining contribution 
$\frac{\alpha_s}{\pi}\left(\delta \mathcal{F}_{\mathrm{rec,qq}}+\delta \mathcal{F}_{\mathrm{hc,qq}}\right)$. 
In analogy with the extraction of $C^{(1)}_{\mathrm{hc},qq}$, this term is obtained by subtracting the counterterm $\mathrm{CT}$ from the first two lines of Eq.~\eqref{eq:FISR-start}, yielding
\begin{equation}
\label{eq:dFISR+CT}
\begin{aligned}
  \frac{\alpha_s(Q)}{\pi}&\left(\delta \mathcal{F}_{\mathrm{rec,qq}}+\delta \mathcal{F}_{\mathrm{hc,qq}}\right)=\int \dZ
  \frac{\alpha_s(Qv^{\frac{1}{a+b_\ell}})}{2\pi} \int_0^{Q^2}\frac{dk_t^2}{k_t^2}
\left(\frac{Q^2v^{\frac{2}{a+b_\ell}}}{k_t^2}\right)^{\varepsilon} \int_0^{2\pi}\frac{d\phi}{2\pi}
\times \\
&\times \left\{\left[\int_{x_\ell}^1 \frac{dz}{z}
\hat P_{qq}(z;\epsilon)
\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)
\frac{f_q(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_q(x_\ell,Qv^{\frac{1}{a+b_\ell}})}+\right.\right. \\
&\left.
  - \int_0^1 dz \frac{2C_\ell}{1-z}
\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)\right] + \\
&-\left[\int_{x_\ell}^1\frac{dz}{z} \hat P_{qq}(z;\epsilon)
\frac{f_q(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_q(x_\ell,Qv^{\frac{1}{a+b_\ell}})}
-\int_0^1dz\frac{2 C_\ell}{1-z}\right] \times\\
&\times \left.\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k)}{v}\right)
\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)
\right\}\,.
\end{aligned}
\end{equation}
To simplify the above expression, we introduce the rescaled variable $
\zeta \equiv V_{\mathrm{sc}}(\{\tilde{p}\},k)/v$, and change variables using the parametrisation of
$V_{\mathrm{sc}}(\{\tilde{p}\},k)$ in Eq.~\eqref{eq:Vsc-param}. This gives
\begin{equation}
	\frac{dk_t^2}{k_t^2}\simeq\frac{2}{\left(a+b_\ell\right)}\frac{d\zeta}{\zeta}\,.
\end{equation}
Using this substitution and taking the limit $\varepsilon \to 0$, Eq.~\eqref{eq:dFISR+CT} simplifies to
\begin{equation}
  \begin{aligned}
    \label{eq:dFrec+dFhq}
\frac{\alpha_s(Q)}{\pi}&\left(\delta \mathcal{F}_{\mathrm{rec,qq}}+\delta \mathcal{F}_{\mathrm{hc,qq}}\right)=\int \dZ \frac{\alpha_s\left(Qv^{1/(a+b_\ell)}\right)}{\pi\left(a+b_\ell\right)}\left\{\int_0^\infty\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}\times\right. \\
&\times \left(\int_{x_\ell}^1 \frac{dz}{z} \hat P_{qq}(z;0)\,\frac{f_q(x_\ell/z,Qv^{1/(a+b_\ell)})}{f_q(x_\ell,Qv^{1/(a+b_\ell)})}
\left[\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)+\right.\right.\\
&\left.-\Theta(1-\zeta)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)\right] +\\
&\left.\left.-\int_0^1 dz \frac{2C_\ell}{1-z}\left[\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right) -\Theta(1-\zeta)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)\right]\right)\right\}\,.
\end{aligned}
\end{equation}
We observe that the second and third line give two contributions that
are separately finite in the collinear limit $\zeta\to 0$. However, in
order to take the limit
$z\to 1$ and obtain a finite results, one needs to compute the expression in
eq.~\eqref{eq:dFrec+dFhq} as a whole. A further simplification can be achieved by adding and subtracting the following term
\begin{multline}
  \label{eq:dFhc-sub}
  \int \dZ \frac{\alpha_s\left(Q v^{1/(a+b_\ell)}\right)}{\pi\left(a+b_\ell\right)}\int_0^\infty\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}
 \int_{x_\ell}^1 \frac{dz}{z}  \hat P_{qq}(z;0)\, \frac{f_q(x_\ell/z,Qv^{1/(a+b_\ell)})}{f_q(x_\ell,Qv^{1/(a+b_\ell)})}\times\\
 \times\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)\,.
\end{multline}
This gives two contributions, namely $\delta \mathcal{F}_{\mathrm{rec,qq}}$ and $\delta \mathcal{F}_{\mathrm{hc,qq}}$:
\begin{equation}
\label{eq:dFrec-qq}
\begin{aligned}
\delta \mathcal{F}_{\mathrm{rec,qq}} = \frac{\alpha_s\left(Q v^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}\int_0^\infty\frac{d\zeta}{\zeta}&\int_0^{2\pi}\frac{d\phi}{2\pi} \int \dZ\int_{x_\ell}^1 \frac{dz}{z} \hat P_{qq}(z;0)\,\frac{f_q(x_\ell/z,Qv^{1/(a+b_\ell)})}{f_q(x_\ell,Qv^{1/(a+b_\ell)})} \\
\times  
&\left[\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)-\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)\right]\,,
\end{aligned}
\end{equation}
\begin{equation}
\label{eq:dFhc-qq}
\begin{aligned}
\delta \mathcal{F}_{\mathrm{hc,qq}} &= \frac{\alpha_s\left(v^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}\left(\int_{x_\ell}^1 \frac{dz}{z}\frac{f_q(x_\ell/z,Qv^{1/(a+b_\ell)})}{f_q(x_\ell,Qv^{1/(a+b_\ell)})} \left(P_{qq}(z)+\gamma^{(0)}_\ell\delta(1-z)\right)\right)\int_0^\infty\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}\times \\
&\times \int \dZ   \left[\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right) -\Theta(1-\zeta)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)\right]\,.
\end{aligned}
\end{equation}

The contribution of the splitting $g\to q\bar q$ can be obtain in a
similar way, with the major difference that there is no soft limit to
subtract and there are no virtual corrections. \question{Do not understand
why we have neither of these?} This gives three contributions, $C_{\mathrm{hc},qg}^{(1)}$, $\delta\mathcal{F}_{\mathrm{hc},qg}$, $\delta\mathcal{F}_{\mathrm{rec},qg}$:
  \begin{multline}
    \label{eq:C1qq-final}
C^{(1)}_{\mathrm{hc}, qg}=  \int_{x_\ell}^1\frac{dz}{z}\,\frac{f_g(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_q(x_\ell,Qv^{\frac{1}{a+b_\ell}})}\times \\ \times
\left[\frac{2 }{a+b_\ell}\left(b_\ell\left(\frac{\ln(1-z)}{1-z}\right)_+-\frac{1}{(1-z)_+}
        \left(\langle\ln d_\ell g_\ell\rangle-b_\ell \ln\frac{2E_\ell}{Q}\right)\right)(1-z)\hat P_{qg}(z;0)-P_{qg}^{(\varepsilon)}(z)\right]\,.
  \end{multline}
\begin{equation}
\label{eq:dFrec-qg}
\begin{aligned}
\delta \mathcal{F}_{\mathrm{rec},qg} = \frac{\alpha_s\left(Q v^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}\int_0^\infty\frac{d\zeta}{\zeta} &\int_0^{2\pi}\frac{d\phi}{2\pi}\int \dZ\int_{x_\ell}^1 \frac{dz}{z} \hat P_{qg}(z;0)\,\frac{f_g(x_\ell/z,Qv^{1/(a+b_\ell)})}{f_q(x_\ell,Qv^{1/(a+b_\ell)})} \\
&\times  
\left[\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)-\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)\right]\,,
\end{aligned}
\end{equation}
\begin{multline}
\label{eq:dFhc-qg}
\delta \mathcal{F}_{\mathrm{hc},qg} = \frac{\alpha_s\left(Qv^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}\left(\int_{x_\ell}^1 \frac{dz}{z}\frac{f_g(x_\ell/z,Qv^{1/(a+b_\ell)})}{f_q(x_\ell,Qv^{1/(a+b_\ell)})} P_{qg}(z)\right)\int_0^\infty\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}\times \\
\times \int \dZ   \left[\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right) -\Theta(1-\zeta)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)\right]\,.
\end{multline}
Where we have introduced the splitting functions
\begin{equation}
  \label{eq:Pqg}
  \hat P_{qg}(z;\varepsilon) = \hat P_{qg}(z;0)+\varepsilon \hat P_{qg}^{(\varepsilon)}(z)\,,\quad
 \hat P_{qg}(z;0)= P_{qg}(z) = T_R \left[z^2+(1-z)^2\right]\,,
   \quad \hat P_{qg}^{(\varepsilon)}(z) = -2T_R z(1-z)\,.
 \end{equation}
 
 The two channels $q\to qg$ and $g\to q\bar{q}$ are instances of a
general structure. The coefficient $C^{(1)}_{\mathrm{hc}, qg}$ can be written in the general form,
  \begin{equation}
    \label{eq:Cell-ISR}
    C_{\mathrm{hc},\ell}^{(1)} = \sum_j  \int_{x_\ell}^1\frac{dz}{z}\,C_{ij}^{(1)}(z) \frac{f_j(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_i(x_\ell,Qv^{\frac{1}{a+b_\ell}})}\,,
  \end{equation}
  where $i$ is the flavour of leg $\ell$. The coefficients $C_{ij}^{(1)}(z)$ are given in terms of the splitting functions in $4-2\varepsilon$ dimensions
  \begin{equation}
    \label{eq:Pij-general}
    \hat P_{ij}(z;\varepsilon)= \hat P_{ij}(z;0)+\varepsilon P_{ij}^{(\varepsilon)}(z)\,,
  \end{equation}
  and have the following expression
  \begin{multline}
    \label{eq:Cij-ISR}
    C_{ij}^{(1)}(z) = -\hat P_{ij}^{(\varepsilon)}(z) \\ +\frac{2 }{a+b_\ell}\left(b_\ell\left(\frac{\ln(1-z)}{1-z}\right)_+-\frac{1}{(1-z)_+}
      \left(\langle\ln d_\ell g_\ell\rangle-b_\ell \ln\frac{2E_\ell}{Q}\right)\right)(1-z)\hat P_{ij}(z;0)\,.
  \end{multline}
  The explicit expressions of the splitting
  functions are
  \begin{subequations}
  	\begin{align}
	& \hat P_{qq}(z;\varepsilon)=C_F \frac{1+z^2}{1-z}-\varepsilon\,C_F (1-z)\,,\\
	& \hat P_{qg}(z;\varepsilon)=T_R \left[z^2+(1-z)^2\right]-\varepsilon\,T_R\left[2z(1-z)\right]\,, \\
	  \label{eq:Pgq-epsilon}
	& \hat P_{gq}(z;\varepsilon) =C_F \frac{1 + (1-z)^2}{z} - \varepsilon\,C_Fz \,, \\
	 \label{eq:Pgg-epsilon}
	& \hat P_{gg}(z;\varepsilon) =2C_A\left[\frac{z}{1-z} + \frac{1-z}{z} + z(1-z)\right]\,.
	\end{align}
  \end{subequations}
  For the final-state leg ($\ell=3$), $C^{(1)}_{{\rm hc},3}$ depends
  on the flavour of the outgoing parton and can be obtained by
  the replacement
  \begin{equation}
    \label{eq:C-recovery}
    \frac{1}{z}\frac{f_j(x_\ell/z,Qv^{\frac{1}{a+b_\ell}})}{f_i(x_\ell,Qv^{\frac{1}{a+b_\ell}})}\Theta(z-x_\ell)\to 1
  \end{equation}
  and integrating over $z$. This gives exactly
  the same results as in Ref.~\cite{Arpino:2019ozn},
  such that for a quark or antiquark,
  \begin{equation}
    \label{eq:C1hc-q}
    C_{\mathrm{hc},3}^{(1)}=C_F\left(\frac{7}{2}\frac{b_3}{a+b_3}+\frac{3}{a+b_3}\left(\ln\langle d_3 g_3\rangle-b_3\ln\frac{2E_3}{Q}\right)+\frac 12\right)\,,
  \end{equation}
  while for a gluon
  \begin{equation}
    \label{eq:C1hc-q}
    \begin{split}
    C_{\mathrm{hc},3}^{(1)}& =\left(\frac{67}{18}C_A-\frac{13}{9}T_R n_f\right)\frac{b_3}{a+b_3}\\ & +\frac{1}{a+b_3}\left(\frac{11}{3}C_A-\frac{4}{3}T_R n_f\right)\left(\ln\langle d_3 g_3\rangle-b_3\ln\frac{2E_3}{Q}\right)+\frac{1}{3}T_R n_f\,.
    \end{split}
  \end{equation}

Similarly, we find that $\delta \mathcal{F}_{\rm hc}$ can be written generally as
 \begin{multline}
   \label{eq:dFhc-ISR}
   \delta\mathcal{F}_{\rm hc}(\lambda)=\sum_\ell \frac{\alpha_s\left(Qv^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}B_\ell \int_0^{\infty}\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}
 \int \dZ  \times\\ \times  \left[\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right) -\Theta(1-\zeta)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)\right]\,.
 \end{multline}
 For initial-state radiation, $B_\ell$ is given by
 \begin{equation}
   \label{eq:Bell-ISR}
B_\ell=\sum_j\int_{x_{\ell}}^1 \frac{dz}{z} \left(P_{ij}(z)-\gamma_\ell^{(0)}\delta_{ij}\right) \frac{f_j(x_{\ell}/z,Qv^{1/(a+b_\ell)})}{f_i(x_\ell,Qv^{1/(a+b_\ell)})}  \,,
 \end{equation}
 where $P_{ij}(z)$ are the full Altarelli-Parisi splitting functions,
 \begin{subequations}
 \label{eq:full-AP}
\begin{align}
	&P_{qq}(z)=C_F\left[\frac{1+z^2}{1-z}\right]_+\,,\\
	&P_{qg}(z)=T_R\left[z^2+(1-z)^2\right]\,,\\
	&P_{gq}(z)=C_F\left[\frac{1+(1-z)^2}{z}\right]\,,\\
	&P_{gg}(z)=2C_A\left[\frac{z}{(1-z)_+}+\frac{1-z}{z}+z(1-z)\right]+2\pi\beta_0\delta(1-z)\,.
\end{align}
 \end{subequations}
  For final-state radiation, the same substitution can be
  used as in Eq.~\eqref{eq:C-recovery}, such that $B_\ell=\gamma_\ell^{(0)}$,
 where 
 \begin{equation}
  \label{eq:gammal-flav}
  \gamma_\ell^{(0)} = \left\{
    \begin{split}
      & \frac{3}{2}C_F\,,\qquad\text{if $\ell$ is a quark,} \\
      & 2\pi \beta_0 \,,\qquad\text{if $\ell$ is a gluon}\,.
    \end{split}
    \right.
\end{equation}

The recoil correction, $\delta\mathcal{F}_{\rm rec}$ can be written in the general form,
 \begin{multline}
   \label{eq:dFrec-ISR}
   \delta\mathcal{F}_{\rm rec}(\lambda)=\sum_\ell \frac{\alpha_s\left(Qv^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}\int_0^1 dz P_\ell(z) \int_0^{\infty}\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}
 \int \dZ  \times\\ \times  \left[\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right) -\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)\right]\,.
\end{multline}
For an initial-state leg,
\begin{equation}
  \label{eq:Pell-ISR}
  P_\ell(z)=\sum_j \hat P_{ij}(z;0) \frac{f_j(x_\ell/z,Qv^{1/(a+b_\ell)})}{z f_i(x_\ell,Qv^{1/(a+b_\ell)})} \Theta(z-x_\ell)\,,
\end{equation}
where $ \hat P_{ij}(z;0)$ are defined in Eq.~\eqref{eq:Pij-general}.
For a final-state leg, 
\begin{equation}
  \label{eq:Pell-FSR}
  P_\ell(z)=\left\{
    \begin{split}
      &C_F \frac{1+z^2}{1-z}\,,\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad\text{if $\ell$ is a quark,}\\
      &C_A\left(\frac{2z}{1-z}+z(1-z)\right)+T_R\left[z^2+(1-z)^2\right]\,,\qquad\text{if $\ell$ is a gluon}\,.
    \end{split}
  \right.
\end{equation}
\question{I still don't understand how we get this?}
\ABquestion{I think the answer for the gluon is incorrect as it does not 
integrate to the correct result.}

 
\subsection{Gluon entering the hard scattering}
\label{sec:gluon-in}

We now consider the case in which a gluon enters the hard scattering.
Two collinear splittings contribute: an incoming quark emitting a
collinear gluon ($q\to gq$), illustrated in
Fig.~\ref{fig:g-incoming-qgq}, and an incoming gluon undergoing a 
splitting into two gluons ($g\to gg$), illustrated in Fig.~\ref{fig:g-incoming-ggg}.
These give the initial-state collinear contributions to the $qg/gq$
subprocesses for $Z$+jet production.

\begin{figure}[H]
\centering

\begin{subfigure}{0.48\textwidth}
\centering
\begin{tikzpicture}[scale=0.85]
\begin{feynman}
  \vertex (H) at (0,0);
  \vertex (v1) at (-2.0, 1.5);
  \vertex (q0) at (-4.2, 3.0);
  \vertex (qbar) at (-4.2, -2.2);
  \vertex (gISR) at (0.6, 2.95);
  \vertex (Z)   at (3.8,  1.6);
  \vertex (gout) at (3.8, -1.6);
  \diagram*{
    (q0)   -- [fermion] (v1),
    (v1)   -- [gluon,   edge label={$z$}] (H),
    (v1)   -- [fermion]                   (gISR),
    (qbar) -- [fermion]              (H),
    (H)    -- [boson]                     (Z),
    (H)    -- [fermion]                     (gout),
  };
  \draw[fill=white] (H) circle (0.8cm) node {$\mathcal{H}_{gq}$};
\end{feynman}
\node[above left]  at (-4.2, 3.0)  {$q$};
\node[above right] at (-2.1, 2.3)  {$1-z$};
\node[below left]  at (-4.2, -2.2) {$q$};
\node[above right] at (3.8,  1.6)  {$Z$};
\node[below right] at (3.8, -1.6)  {$q$};
\end{tikzpicture}
\caption{$q \to gq$ splitting}
\label{fig:g-incoming-qgq}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\begin{tikzpicture}[scale=0.85]
\begin{feynman}
  \vertex (H) at (0,0);
  \vertex (v1) at (-2.0, 1.5);
  \vertex (q0) at (-4.2, 3.0);
  \vertex (qbar) at (-4.2, -2.2);
  \vertex (gISR) at (0.6, 2.95);
  \vertex (Z)   at (3.8,  1.6);
  \vertex (gout) at (3.8, -1.6);
  \diagram*{
    (q0)   -- [gluon] (v1),
    (v1)   -- [gluon, edge label={$z$}] (H),
    (v1)   -- [gluon]                   (gISR),
    (qbar) -- [fermion]            (H),
    (H)    -- [boson]                   (Z),
    (H)    -- [fermion]                   (gout),
  };
  \draw[fill=white] (H) circle (0.8cm) node {$\mathcal{H}_{gq}$};
\end{feynman}
\node[above left]  at (-4.2, 3.0)  {$g$};
\node[above right] at (-2.1, 2.3)  {$1-z$};
\node[below left]  at (-4.2, -2.2) {$q$};
\node[above right] at (3.8,  1.6)  {$Z$};
\node[below right] at (3.8, -1.6)  {$q$};
\end{tikzpicture}
\caption{$g \to gg$ splitting}
\label{fig:g-incoming-ggg}
\end{subfigure}

\caption{Initial-state radiation contributions to the $gq$ channel of $Z$+jet production, from the $q \to gq$ splitting (left)and the $g \to gg$ splitting (right).  The hard scattering is denoted by $\mathcal{H}_{gq}$, where the subscripts indicate the incoming partons.
}
\label{fig:g-incoming}
\end{figure}

For an incoming gluon, the derivation of
$C_{{\rm hc},\ell}^{(1)}$, $\delta\mathcal{F}_{\rm hc}$, and
$\delta\mathcal{F}_{\rm rec}$ follows the same steps as
Sec.~\ref{sec:quark-in}, resulting in the expressions found
in Eqs.~\eqref{eq:Cell-ISR},~\eqref{eq:dFhc-ISR}, and~\eqref{eq:dFrec-ISR}
respectively. 

However, a gluon entering the hard scattering introduces an additional
contribution that has no counterpart in the quark case: spin
correlations. \todo{Explain why they arise here.} These correlations give rise
to an additional NNLL contribution, $\Delta\mathcal{F}_{\rm rec}$.

To account for this, we introduce the spin-density matrix for the hard process
in $4-2\varepsilon$ dimensions. Neglecting terms that are antisymmetric in $\mu,\nu$, the general form is \ABquestion{The $P_{L}$ pre-factors?}
\begin{equation}
\rho^{\mu\nu} = \frac{d^{\mu\nu} }{2(1-\varepsilon)}
+ \mathcal{T}_\ell(\{p\};\varepsilon)\left(2n_{\rm in}^\mu n_{\rm in }^\nu - \frac{d^{\mu\nu} }{1-\varepsilon}\right) \,,
\end{equation}
where we choose $d^{\mu\nu}$ to be the sum over physical polarisations transverse to both $p_1$ and $p_2$,
\begin{equation}
d^{\mu\nu} = -g^{\mu\nu} + \frac{p_1^\mu p_2^\nu + p_2^\mu p_1^\nu}{p_1 \cdot p_2}\,,
\end{equation}
and $n_{\rm in}$ is a space-like unit vector that lies in the plane identified
by the beam and the $Z$ boson, satisfying $n_{\rm in}^2=-1$ and $n_{\rm in}\cdot p_1=n_{\rm in}\cdot p_2=0$. We obtain the following relations: 
\begin{equation}
  \begin{split}
  &d^{\mu\nu}d_{\mu\nu}=-g^{\mu\nu}d_{\mu\nu}=2(1-\varepsilon)\,,\\ & d_{\mu\nu} \left(2n_{\rm in}^\mu n_{\rm in }^\nu - \frac{d^{\mu\nu}}{1-\varepsilon}\right)=-g^{\mu\nu}\left(2 n_{\rm in}^\mu n_{\rm in }^\nu - \frac{d^{\mu\nu}}{1-\varepsilon}\right)=0\,,\\ &
  \left(2 n_{\rm in}^\mu n_{\rm in }^\nu - \frac{d^{\mu\nu}}{1-\varepsilon}\right)\left(2 n_{{\rm in},\mu} n_{{\rm in },\nu} - \frac{d^{\mu\nu}}{1-\varepsilon}\right) = 2 \frac{1-2\varepsilon}{1-\varepsilon}\,.
  \end{split}
\end{equation}
This implies
\begin{equation}
  \label{eq:PL}
  \mathcal{T}_\ell(\{\tilde p\};\varepsilon) = \frac{1-\varepsilon}{2(1-2\varepsilon)} \left(2 n_{\rm in}^\mu n_{\rm in }^\nu - \frac{d^{\mu\nu}}{1-\varepsilon}\right) \rho_{\mu\nu}\,.
\end{equation}
%For the $gq \to Zq$ channel, the polarization factors read
%\begin{equation}
%\begin{aligned}
%\mathcal{P}_L &= \frac{2su}{s^2 + u^2 + 2M^2 t}\,,\\
%\mathcal{P}_C &= \frac{g_V g_A}{g_V^2 + g_A^2}\frac{s^2 - u^2}{s^2 + u^2 + 2M^2 t}\,.
%\end{aligned}
%\end{equation}

The collinear splitting probability is then obtained by contracting
$\rho^{\mu\nu}$ with the unaveraged splitting functions
\begin{equation}
\left[\hat{P}_{ga}^{(1)}(z,\vec k_t;\varepsilon)\right]^{\mu\nu}
= -g^{\mu\nu}\,\hat{P}_{ga}(z;\varepsilon)
+ \left(\frac{g^{\mu\nu}}{1-\varepsilon} + 2\frac{k_\perp^\mu k_\perp^\nu}{k_t^2}\right)
2\hat G_{ga}(z;\varepsilon)\,, \qquad a = g,q\,,
\end{equation}
with $k_\perp\cdot k_\perp = -k_t^2$, and
\begin{equation}
  \label{eq:Grho}
  d_{\mu\nu}\left(\frac{g^{\mu\nu}}{1-\varepsilon} + 2\frac{k_\perp^\mu k_\perp^\nu}{k_t^2}\right)=0\,.
\end{equation}
The azimuthally averaged splitting functions $\hat{P}_{ga}(z;\varepsilon)$ with $a=q,g$ are those present in Eqs.~\eqref{eq:Pgq-epsilon} and.~\eqref{eq:Pgg-epsilon} respectively, and the
spin-correlated functions are \ABquestion{We need to check this.}
\begin{equation}
\begin{aligned}
\hat G_{gg}(z) &= -(1-\varepsilon)C_A \frac{1-z}{z}\,,\qquad
\hat G_{gq}(z) &= C_F \frac{1-z}{z}\,.
\end{aligned}
\end{equation}
Performing the contraction, using the fact $k_\perp \cdot n_{\rm in }=-k_t \cos\phi$ we obtain
\begin{equation}
\rho^{\mu\nu}\left[\hat{P}_{ga}(z,\vec k_t;\varepsilon)\right]_{\mu\nu}
= \hat{P}_{ga}(z;\varepsilon)
+  \mathcal{T}_\ell(\{\tilde p\};\varepsilon)\,\hat G_{ga}(z;\varepsilon)\left[2(1-\varepsilon)\cos^2\phi - 1\right]\,,
\qquad a = g,q\,.
\end{equation}
The above equation gives the probability for a splitting $a\to g
a$. For $\hat{P}_{ga}(z;\varepsilon)$, we can repeat exactly the same
analysis as in section~\ref{sec:quark-in}, and obtain identical
expressions for $C_{\mathrm{hc},\ell}^{(1)}$,
$\delta\mathcal{F}_{\rm hc}$ and $\delta\mathcal{F}_{\rm rec}$. The
functions $\hat{G}_{ga}(z;\varepsilon)$ can be handled using the same
strategy as in~\cite{Arpino:2019ozn}, and yields the NNLL contribution
\ABquestion{Do not understand the structure of 
the expression here? Is the total $\Delta\mathcal{F}_{\rm rec}$ 
here the sum over legs of this with the $\alpha_s(Q)/\pi$?}
\begin{equation}
\begin{aligned}
\label{eq:DFrec-ISR}
  \Delta&\mathcal{F}^{(\ell)}_{\mathrm{rec}}(\lambda) =
\int_0^\infty \frac{d\zeta}{\zeta}
\int_0^\pi \frac{d\phi}{\pi}\left[2\cos^2\phi - 1\right]
\int_0^1\!dz\,
\Delta P_g(z)
\\
&\quad \times \int \dZ \left[
\Theta\left(1 - \frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)
- \Theta(1-\zeta)\,
\Theta\left(1 - \frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)
\right]\,,
\end{aligned}
\end{equation}
where for initial-state radiation
\begin{equation}
\label{eq:DPg-ISR}
\Delta P_g(z) =
4\left(\frac{1-z}{z}\right)\frac{\Theta(z-x_\ell)}{z f_g(x_\ell,Q v^{1/(a+b_\ell)})}
\left[C_A f_g\left(\frac{x_\ell}z,Q v^{1/(a+b_\ell)}\right) + C_F f_q\left(\frac{x_\ell}z,Q v^{1/(a+b_\ell)}\right)\right]\,,
\end{equation}
while for final-state radiation
\begin{equation}
\label{eq:DPg-FSR}
\Delta P_g(z) =4z(1-z)\left[\frac{C_A}{2}-T_Rn_f\right]\,.
\end{equation}

\section{ NNLL resummation: one-jettiness}
\label{sec:one-jettiness}

One-jettiness is an rIRC-safe, continuously global observable. It vanishes
at Born level,
\[
\tau_1(\{p_1,p_2,p_3\})
=
\tau_1(\{\tilde p_1,\tilde p_2,\tilde p_3\})
=0,
\]
and satisfies the additional scaling properties for an arbitrary number of soft and/or collinear
emissions discussed in Sec.~\ref{sec:obs-req}. It therefore lies within the
class of observables to which the \textsc{ARES} formalism applies.

We begin by considering the behaviour of the observable in the presence of a
single soft emission collinear to leg $\ell$. By definition,
 $k\cdot p_\ell/E_\ell=k_t^{(\ell)} e^{-\eta^{(\ell)}}$, where
  $k_t^{(\ell)}$ is the relative transverse momentum with respect to
  the emitting leg $\ell$ and $\eta^{(\ell)}$ its rapidity with respect to the same leg. 
  Using $\tau_1 = \mathcal{T}_1/M$ with $\mathcal{T}_1$ defined in
Eq.~\eqref{eq:T1}, the soft-collinear limit takes the same form for every leg:
    \begin{equation}
    \label{eq:tau1-sc-1emsn}
    \tau_1(\{\tilde p\},k)\simeq \tau^{\rm sc}_1(\{\tilde p\},k) =  \frac{2E_\ell}{Q_\ell}\frac{k_t^{(\ell)}}{M} e^{-\eta^{(\ell)}}\,.
  \end{equation}
  From eq.~\eqref{eq:tau1-sc-1emsn} we read off the single-emission
parameters,
\begin{equation}
  \label{eq:tau1-1em-props}
  a=1\,, \quad b_\ell=1\,, \quad d_\ell=\frac{Q}{M}\frac{2E_{\ell}}{Q_{\ell}} \,, \quad g_\ell(\phi) = 1\,, \qquad \ell=1,2,3\,.
\end{equation}

The single-emission parameters in Eq.~\eqref{eq:tau1-1em-props} are sufficient
to determine $H_{ij}^{(1)},R_{\rm NNLL}$ and $C_{\mathrm{hc},\ell}^{(1)}$ within  Eq.~\eqref{eq:Sigma-NNLL}  by directly substituting these parameters into the corresponding general expressions \add{add in equations}. Similarly, since one-jettiness is an additive observable, the NLL multiple-emission function $\mathcal{F}_{\rm NLL}$ is given by the
additive result in Eq.~\eqref{eq:FNLL-add-final}.

The remaining task is therefore to compute the genuine NNLL multiple-emission
corrections entering the cumulant through Eq.~\eqref{eq:dFNNLL}. These
corrections arise from different kinematic regions of phase space, and we
organise their calculation according the kinematic region from which each contribution arises:
soft-collinear, hard-collinear, and soft wide-angle radiation.


Since one-jettiness is additive, these NNLL corrections take a particularly
simple form. The corresponding general additive expressions are collected in
Appendix~\ref{ap:add-res-NNLL}. In the following subsections, we briefly
state the general equation of each correction to and then evaluate it for
one-jettiness.


\subsection{Soft and collinear contributions}
\label{sec:sc-contributions}

The NNLL corrections associated with the soft-collinear region are
$\delta \mathcal{F}_{\rm hc}$, $\delta \mathcal{F}_{\rm sc}$, and
$\delta \mathcal{F}_{\rm correl}$. All three contributions make use of the soft-collinear limit defined in Eq.~\eqref{eq:Vsc}. On top of the soft-collinear
emissions $k_1,\dots,k_n$ widely separated in angle, which is already present in the
NLL resummation, the corrections $\delta \mathcal{F}_{\rm hc}$, $\delta \mathcal{F}_{\rm sc}$ involve an additional soft-collinear emission $k$. This emission is parametrised in terms of the usual variable $\zeta=\Vsc{k}/v$, the emitting leg $\ell$, and its azimuthal
angle $\phi^{(\ell)}$. \ABquestion{Unsure on the limits used making sense as this parameterisation
$\zeta=\Vsc{k}/v$?}

\ndp{The correlated correction, $\delta \mathcal{F}_{\rm correl}$ instead
accounts for correlated double emissions. 
In this case the additional soft-collinear emission splits into 
two partons, $k_a$ and $k_b$, with the splitting described by 
the variables introduced in Eq.~\eqref{eq:dFcorrel-general}.
Therefore, while $\delta\mathcal{F}_{\rm hc}$ and $\delta\mathcal{F}_{\rm sc}$
describe configurations involving independent emissions widely separated in angle,
$\delta\mathcal{F}_{\rm correl}$ captures the effect of correlated
splittings of a single soft-collinear emission into a $q\bar q$ pair or a
gluon pair.}

The general expression for $\delta \mathcal{F}_{\rm hc}$ is given by 
\begin{multline}
  \label{eq:dFhc-general}
   \delta\mathcal{F}_{\rm hc}=\sum_\ell \frac{\alpha_s\left(Qv^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}B_\ell \int_0^{\infty}\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}
 \int \dZ  \times\\ \times  \left[\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right) -\Theta(1-\zeta)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},\{k_i\})}{v}\right)\right]\,.
\end{multline}

For an additive observable, this expression can be simplified using the
additive form given in Eq.~\eqref{eq:dFhc-general-add}. Substituting the
single-emission parameters for one-jettiness from
Eq.~\eqref{eq:tau1-1em-props} gives
 \begin{equation}
  \label{eq:dFhc-tau_1}
    \delta \mathcal{F}_{\rm hc}(\lambda) = \frac{\alpha_s(\sqrt{\tau_1}\,Q)}{2\,\alpha_s(Q)}\frac{d\FNLL}{d\RpNLL} \sum_{\ell}B_\ell\,.
  \end{equation}
  
  For initial-state legs, the $z$-integration entering $B_\ell$
(see Eq.~\eqref{eq:Bell-ISR}) cannot be performed analytically due to the
presence of PDFs. For the final-state leg,
$B_3=-\gamma_3^{(0)}$, reproducing the corresponding result for
three-jet production in $e^+e^-$ annihilation~\cite{Arpino:2019ozn}.

The second soft-collinear contribution is
$\delta\mathcal{F}_{\rm sc}$, whose general expression is
\begin{multline}
	\label{eq:dFsc-general}
	\delta\mathcal{F}_{\mathrm{sc}}(\lambda)=\frac{\pi}{\alpha_s(Q)}\int_0^\infty\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}\sum_{\ell=1}^3\left[\Delta R'_{\ell,\mathrm{NNLL}}+R''_{\ell,\mathrm{NNLL}}\left(\ln\frac{d_\ell g_\ell(\phi)}{\zeta}-b_\ell\ln\frac{2E_\ell}{Q}\right)\right]\times 
	 \\ \times \int \dZ  \left[\Theta\left(1-\frac{\Vsc{k,\{k_i\}}}{v}\right)
       -\Theta\left(1-\frac{\Vsc{\{k_i\}}}{v}\right)\Theta(1-\zeta)
     \right]\,.
\end{multline}

 Using the additive result in Eq.~\eqref{eq:dFsc-general-add}, the result for the one-jettiness is
\begin{multline}
\delta\mathcal{F}_{\mathrm{sc}}(\lambda)=-\FNLL(\lambda)\frac{\pi}{\alpha_s(Q)}\sum_{\ell=1}^3\left[\left(\Delta R'_{\ell,\mathrm{NNLL}}+ R''_{\ell,\mathrm{NNLL}}\ln\frac{Q^2}{M Q_{\ell}}\right)\left(\psi^{(0)}\left(1+\RpNLL\right)+\gamma_E\right)\right. \times\\
 \left.+\frac{R''_{\ell,\mathrm{NNLL}}}{2}\left(\left(\psi^{(0)}(1+\RpNLL)+\gamma_E\right)^2-\psi^{(1)}(1+\RpNLL)+\frac{\pi^2}{6}\right)\right]\,.
\end{multline}


The final soft-collinear contribution is the correlated correction
$\delta\mathcal{F}_{\rm correl}$. Its general expression is
\begin{equation}
\begin{aligned}
\label{eq:dFcorrel-general}
\delta&\mathcal{F}_{\mathrm{correl}}(\lambda)=\sum_{\ell=1}^3\frac{\lambda R''_{\ell,\mathrm{NNLL}}}{2a\beta_0\alpha_s(Q)}
\int_0^\infty\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi^{(\ell)}}{2\pi}\int_0^\infty
\frac{d\mu^2}{\mu^2(1+\mu^2)}\int_0^1dz\int_0^{2\pi}\frac{d\phi}{2\pi}\frac{1}{2!}\mathcal{A}^2(z,\mu,\phi)\times\\
&\int\dZ\left[ \Theta\left(1-\frac{V_{\mathrm{sc}}\left(\{\tilde{p}\},k_a,k_b,\{k_i\}\right)}{v}\right) - 
\lim_{\mu^2\to0}\Theta\left(1-\frac{V_{\mathrm{sc}}\left(\{\tilde{p}\},k_a+k_b,\{k_i\}\right)}{v}\right) \right]\,,
\end{aligned}
\end{equation}
where $k=k_a+k_b$, $\mu^2=m^2/k_t^2$ is the rescaled invariant mass of the
splitting, and $\mathcal{A}^2$ is the double-emission function defined in
Ref.~\cite{Arpino:2019ozn}. \ABquestion{Is it necessary for me to
explain these any further?}

Since one-jettiness is formally equivalent to thrust in $e^+e^-$ annihilation, we adapt the
angularity result of Ref.~\cite{Banfi:2018mcq} at
$x=0$:\todo{Explain more explicitly why one-jettiness is like thrust here.}
\begin{equation}
 	f^{(\ell)}_{\rm correl}\left(z,\mu,\phi,\phi^{(\ell)}\right)=\sqrt{1+\mu^2}\,,
	\label{eq: f-correl-tau_1}
 \end{equation}
giving
\begin{equation}
\label{eq:dFcorrel-tau_1}
	\delta  \mathcal{F}_{\rm correl}(\lambda)=\frac{\pi}{2\alpha_s(Q)}\lambda\,R'' \zeta_2 \FNLL(\lambda)\,.
\end{equation}

\subsection {Hard and collinear contributions}
\label{sec:hc-contributions}

We next consider the NNLL corrections originating from the hard-collinear
region. \ndp{These corrections arise from the fact that a hard-collinear emission
modifies the relation between the Born kinematics and the ensemble of
soft-collinear emissions. }
They are separated into two recoil contributions:
$\delta\mathcal{F}_{\rm rec}$, which arises from a hard-collinear emission 
recoiling against the soft-collinear ensemble at the level of the kinematics, 
and $\Delta\mathcal{F}_{\rm rec}$, which accounts for the correlations between the
spin of a hard-collinear gluon and the event plane. The latter contribution first appears for near-to-planar three-jet configurations~\cite{Arpino:2019ozn}.
Both corrections involve, in addition to the soft-collinear limit of Eq.~\eqref{eq:Vsc}, \ABquestion{Is it using this s-c limit?} the hard-collinear limit
\begin{equation}
  \label{eq:Vhc}
  \frac{V_{\mathrm{hc}}(\{\tilde{p}\}, k_{\mathrm{hc}}, k_1,\dots,k_n)}{v}\equiv \lim_{v\to 0}\frac{\Vfull{k_{\mathrm{hc}}, k_1,\dots,k_n}}{v}\,,
\end{equation}
where $k_{\mathrm{hc}}$ is a single hard-collinear parton 
while $k_1,\cdots k_n$ represent an ensemble independently emitted 
 soft and collinear emissions.

The recoil correction $\delta\mathcal{F}_{\rm rec}$ is described by the
general expression
 \begin{multline}
 \label{eq:dFrec-general}
   \delta\mathcal{F}_{\rm rec}=\sum_\ell \frac{\alpha_s\left(Qv^{1/(a+b_\ell)}\right)}{\alpha_s(Q)\left(a+b_\ell\right)}\int_0^1 dz\,P_\ell(z) \int\frac{d\zeta}{\zeta}\int_0^{2\pi}\frac{d\phi}{2\pi}
 \int \dZ  \times\\ \times  \left[\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k_{\mathrm{hc}},\{k_i\})}{v}\right) -\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)\right]\,.
\end{multline}

For additive observables, the hard-collinear and soft-collinear limits are related by
\ABquestion{Is this just for additive observables and are limits the correct word
to use?}
\begin{equation}
V_{\mathrm{hc}}(\{\tilde{p}\},k_{\rm hc})
=
f_{\mathrm{hc}}(z)
V_{\mathrm{sc}}(\{\tilde{p}\},k),
\end{equation}
and the corresponding additive expression in
Eq.~\eqref{eq:dFrec-general-add} can be used. For one-jettiness, 
this is \ABquestion{Cannot see how we get it from Ref.~\cite{Banfi:2018mcq} at $x=0$?}
\begin{equation}
\label{eq:tau_1-fhc}
f_{\mathrm{hc}}(z)=
\begin{cases}
z\,, \quad \ell=1,2\,, \\
1/z\,, \quad \ell=3\,.
\end{cases}
\end{equation}
\add{Can we write why we have different for ISR and FSR?}

For the initial-state legs, the $z$-integration cannot be performed
analytically because the splitting functions $P_\ell(z)$ involve PDFs (Eq.~\eqref{eq:Pell-ISR}). For the final-state leg $\ell=3$, the result coincides with the corresponding $e^+e^-$ three-jet
calculation. In particular, for a final-state quark one obtains
\begin{equation}
  \label{eq:dFrec-tau_1q}
   \int_0^1dz \, P_\ell(z) \ln f_{\mathrm{hc}}
   =-C_F\left(\frac{5}{4}-\frac{\pi^2}{3}\right)\,,
\end{equation}
while for a final-state gluon
\begin{equation}
  \label{eq:dFrec-tau_1g}
 \int_0^1dz \, P_\ell(z) \ln f_{\mathrm{hc}}
 =-\left(
 C_A\left(\frac{67}{36}-\frac{\pi^2}{3}\right)
 -T_Rn_f\frac{13}{18} \right)\,.
\end{equation}

The second recoil contribution, $\Delta\mathcal{F}_{\rm rec}$, has the
general expression
\begin{multline}
\label{eq:DFrec-general}
\Delta\mathcal{F}_{\mathrm{rec}}(\lambda)=
\frac{\alpha_s\left(Qv^{1/(a+b_g)}\right)}{\left(a+b_g\right)\alpha_s(Q)}\int_0^\infty\frac{d\zeta}{\zeta}
\int_0^{2\pi}\frac{d\phi}{2\pi}\int_0^1dz\Delta P_g\left(z,\phi\right)\int\dZ\times\\ 
\times \left[\Theta\left(1-\frac{V_{\mathrm{hc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right) -\Theta\left(1-\zeta\right)\Theta\left(1-\frac{V_{\mathrm{sc}}(\{\tilde{p}\},k,\{k_i\})}{v}\right)\right]\,.
\end{multline}
Since $\tau_1$ has no azimuthal dependence, the $\phi$-integral over $\Delta P_g(z,\phi)$ vanishes. Using the additive form in Eq.~\eqref{eq:DFrec-general-add}, this gives 
\begin{equation}
  \Delta\mathcal{F}_{\rm rec}(\lambda) = 0\,.
\end{equation}

\subsection{Soft wide angle contributions}
\label{sec:wa-contributions}

The final class of NNLL corrections originates from soft radiation emitted at
wide angles. These comprise the correction
$\delta\mathcal{F}_{\rm wa}$, which accounts for the difference between the
exact observable and its soft-collinear approximation for a single soft
wide-angle emission in the presence of an ensemble of soft-collinear
emissions, and the correction $\Delta\mathcal{F}_{\rm wa}$, which arises from resolved soft wide-angle radiation at NNLL. The latter contribution first appears for near-to-planar three-jet configurations~\cite{Arpino:2019ozn}.

The evaluation of $\delta\mathcal{F}_{\rm wa}$ is naturally expressed in
terms of dipoles $(\ell\bar\ell)$. Therefore, a single emission $k$ can be decomposed along any pair of light-like momenta $(p_{\ell},p_{\bar \ell})$, which constitute the $(\ell \bar\ell)$ dipole, according to 
\begin{equation}
\label{eq:Sudakov-decomp}
k^{(\ell\bar\ell)}=z^{(\ell)}p_{\ell}+z^{(\bar\ell)}p_{\bar\ell}+\kappa_{\ell\bar\ell}\cos\phi^{(\ell\bar\ell)}n_{\rm{in}}^{(\ell\bar\ell)}+\kappa_{\ell\bar\ell}\sin\phi^{(\ell\bar\ell)}n_{\rm{out}}^{(\ell\bar\ell)}\,,
\end{equation}
where
\begin{equation}
	n_{\rm in}^{(\ell\bar\ell)}=\left(\cot\frac{\theta_{\ell\bar\ell}}{2},\frac{\vec{n}_{\ell}+\vec{n}_{\bar\ell}}{\sin\theta_{\ell\bar\ell}}\right)\,,\quad n_{\rm out}^{(\ell\bar\ell)}=\left(0,\frac{n_{\ell}\times n_{\bar\ell}}{\sin\theta_{\ell\bar\ell}}\right)\,,
	\end{equation}
with $\vec n_{\ell/\bar\ell}=\vec p_{\ell/\bar\ell}/E_{\ell/\bar\ell}$.
The remaining variables entering this decomposition are defined in
Sec.~\ref{subsubsec:Rs}.

Unlike the previous corrections, $\delta\mathcal{F}_{\rm wa}$ is expressed in
terms of the soft wide-angle limit of the observable,
\ABquestion{How can we define this limit like the others?}
where a single soft gluon is emitted at large angles.

The general expression for $\delta\mathcal{F}_{\rm wa}$ is:
\begin{equation}
\label{eq:dFwa-general}
\begin{aligned}
	&\delta\mathcal{F}_{\mathrm{wa}}=\frac{\alpha_s(v^{1/a}Q)}{a\,\alpha_s(Q)}2\sum_{(\ell\bar\ell)}C_{\ell\bar\ell}\int_0^\infty \frac{d\zeta}{\zeta}
	\int_{-\infty}^\infty d\eta^{(\ell\bar\ell)}\int_0^{2\pi}\frac{d\phi^{(\ell\bar\ell)}}{2\pi}\times\\
	&\times \int \dZ\left[\Theta\left(1-\frac{V_{\mathrm{wa}}\left(\{\tilde{p}\}, k^{(\ell\bar\ell)},\{k_i\}\right)}{v}\right) -\Theta\left(1-\frac{V_{\mathrm{sc}}\left(\{\tilde{p}\}, k^{(\ell\bar\ell)},\{k_i\}\right)}{v}\right)\right]\,,
\end{aligned}
\end{equation}
where the sum runs over all the dipoles $(\ell\bar\ell)$
that build up the Born event. 

For additive observables, Eq.~\eqref{eq:dFwa-general} simplifies to the form
given in Eq.~\eqref{eq:dFwa-general-add}. The corresponding functions for one-jettiness are
\begin{equation}
\begin{aligned}
	&f_{\rm sc}^{(\ell\bar\ell)}(\eta^{(\ell\bar\ell)},\phi^{(\ell\bar\ell)})=\frac{Q_{\ell\bar\ell}^2}{MQ_{\ell}}e^{-\eta^{(\ell\bar\ell)}}\Theta\left(\eta^{(\ell\bar\ell)}\right)+\frac{Q_{\ell\bar\ell}^2}{MQ_{\bar\ell}}e^{\eta^{(\ell\bar\ell)}}\Theta\left(-\eta^{(\ell\bar\ell)}\right)\,,\\
	&f_{\rm wa}^{(\ell\bar\ell)}(\eta^{(\ell\bar\ell)},\phi^{(\ell\bar\ell)}) =
\min\left\{
\frac{Q_{\ell\bar\ell}^2}{MQ_1}\,e^{-\eta^{(\ell\bar\ell)}},\;
\frac{Q_{\ell\bar\ell}^2}{MQ_2}\,e^{+\eta^{(\ell\bar\ell)}},\;
\frac{Q_{\ell\bar\ell}}{M}\!\left(\cosh(\eta^{(\ell\bar\ell)}-\eta_J)
-\cos(\phi^{(\ell\bar\ell)}-\phi_J)\right)
\right\}\,.
\end{aligned}
\end{equation}
Substituting these functions into the additive expression gives
\begin{equation}
    \delta\mathcal{F}_{\mathrm{wa}}(\lambda) =
    \mathcal{F}_{\rm NLL}(\lambda)\,
    \frac{\alpha_s(\tau_1 Q)}{\alpha_s(Q)}
    \sum_{(\ell\bar\ell)}C_{\ell\bar\ell}
    \left[\ln^2\frac{Q_{\bar\ell}}{Q_\ell}
    - 2\!\left(I_{\ell\bar\ell,\hat\ell}
    + I_{\bar\ell\ell,\hat\ell}\right)\right]\,,
\end{equation}
where the $I_{\ell\bar\ell,\hat\ell}$'s are those found in Eq. (51) in Ref.~\cite{Alioli:2023rxx} (and further defined in Ref.~\cite{Jouttenus:2011wh}) and  the index $\hat\ell$ indicates a leg which is neither $\ell$ or $\bar\ell$.

The second soft wide-angle correction,
$\Delta\mathcal{F}_{\rm wa}$, does not take the wide-angle limit
but instead just uses the same as those in $\delta\mathcal{F}_{\rm {sc}}$
\ABquestion{Why is this a wide angle correction then. if it doesn't have the same limits? Do we need these $v\to0$, I don't think so?}
\begin{multline}
\label{eq:DFwa-general}
	\Delta\mathcal{F}_{\mathrm{wa}}=2\sum_{(\ell\bar\ell)}C_{\ell\bar\ell}\frac{\alpha_s(v^{1/a}Q)}{a\,\alpha_s(Q)}\ln\frac{Q_{\ell\bar\ell}}{Q}\int_0^\infty \frac{d\zeta}{\zeta}
	\int_0^{2\pi}\frac{d\phi}{2\pi}\int \dZ\times \\
	\times \left[\Theta\left(1-\lim_{v\to0}\frac{V_{\mathrm{sc}}\left(\{\tilde{p}\}, k,\{k_i\}\right)}{v}\right) -\lim_{v\to0}\Theta\left(1-\frac{V_{\mathrm{sc}}\left(\{\tilde{p}\}, \{k_i\}\right)}{v}\right)\left(1-\zeta\right)\right]\,.
\end{multline}

For additive observables, Eq.~\eqref{eq:DFwa-general} reduces to the form
given in Eq.~\eqref{eq:DFwa-general-add}. Applying this to one-jettiness
yields
 \begin{equation}
  \label{eq:DFwa-tau_one}
    \Delta \mathcal{F}_{\rm wa}(\lambda) =  \frac{\alpha_s(\tau_1\,Q)}{\alpha_s(Q)}\frac{d\FNLL}{d\RpNLL}2\sum_{(\ell\bar\ell)} C_{\ell\bar\ell}\ln\left(\frac{Q_{\ell\bar\ell}}{Q}\right)\,. 
  \end{equation}
  
In the next section, we will provide an analytical validation of these results.
     
 \section{Validation: N-jettiness subtraction}
 \label{subsec:N-jet-SCET-val}
 
 \todo{In general we need to edit this a bit as the wording is still
 a bit off I think -- I also think we are maybe referring to too many equations? We also need to work out a better correspondence since
 we have $|C|^2$ and PDFs floating around in SCET and we
 cannot really see what we are corresponding to in ARES. }
 
In this section we validate the one-jettiness resummation by
comparing the logarithmic coefficients obtained in the \textsc{ARES}
formalism with those derived in the SCET-based $N$-jettiness
subtraction framework of Ref.~\cite{Gaunt:2015pea}.

\ndp{Ref.~\cite{Gaunt:2015pea} presents a subtraction method for
performing NNLO QCD calculations for arbitrary processes, exploiting
the factorisation of the $N$-jettiness distribution in the limit
$\mathcal{T}_N\to0$. The method uses SCET to determine the IR
singular contributions to $N$-jet cross sections analytically and
constructs suitable $\mathcal{T}_N$-subtractions from them. The
singular structure is organised into hard, beam, jet, and soft
functions. The beam function $B_a$ encodes all collinear emissions
from incoming parton $a$ and depends on its flavour $\kappa_a$ and
momentum fraction $x_a$. The jet function $J_c$ encodes collinear
emissions from the outgoing parton $c$ and depends on its flavour
$\kappa_c$. The soft function $S_\kappa$, where
$\kappa\equiv\{\kappa_a,\kappa_b,\kappa_c\}$ denotes the full
partonic channel, describes all soft emissions between the partons.}

Specialising this to the one-jettiness ($N=1$), we find that the singular part of the distribution ($\tau_1\to0$) factorises as 
\begin{equation}
	\label{eq:Cnm-def}
	\frac{\mathrm{d}\sigma^{\rm sing}(\{p\})}{\mathrm{d}\tau_1}
	=\sum_{m\geq0}\left[\mathcal{C}_{-1}^{(m)}(\{p\},M,\mu_R)\,\delta(\tau_1)
	+\sum_{n=0}^{2m-1}\mathcal{C}_n^{(m)}(\{p\},M,\mu_R)\,\mathcal{L}_n(\tau_1)\right]
	\left(\frac{\alpha_s(\mu_R)}{4\pi}\right)^m\,,
\end{equation}
where 
\begin{equation}
	\mathcal{L}_n(\tau_1)=\left[\frac{\theta(\tau_1)\ln^n(\tau_1)}{\tau_1}\right]_+\,.
\end{equation} 
are the standard plus distributions. The subtraction coefficients
$\mathcal{C}_n^{(m)}$ are expressed in terms of the hard, beam, jet,
and soft functions, where their explicit expressions are collected in
Appendix~\ref{ap:N-jet-ing}, which follows the $a,b,c$ notation which has the correspondence with the leg notation used in the \textsc{ARE}
formalism as $\{a,b,c\}\equiv\{\ell=1,\ell=2,\ell=3\}$.
 
Within the \textsc{ARES} formalism, the cumulant expands as
 \begin{equation}
 \label{eq:sig-coeff}
	\Sigma_{ij}(v)=\sum_{n=1}^{\infty}\sum_{m=0}^{n+1} h_{nm}\bar{\alpha}^n_sL^m\,,
\end{equation}
where $\bar{\alpha}_s\equiv\alpha_s(Q)/2\pi$ and the coefficients
$h_{nm}$ are given in Appendix~\ref{ap:ARES-coeff}. To compare the SCET results with the \textsc{ARES} expansion, it is necessary to
establish a correspondence between the two sets of coefficients. Since the \textsc{ARES}
coefficients arise from the integrated distribution while those of
Ref.~\cite{Gaunt:2015pea} are differential, and accounting for the difference between
the $4\pi$ and $2\pi$ conventions used for $\alpha_s$, we find
\begin{equation}
\label{eq:coeff-convent}
h_{mn}=\frac{\mathcal{C}_{n-1}^{(m)}}{n}\,\frac{(-1)^n}{2m}.
\end{equation}
For the $\delta(\tau_1)$ coefficient,
$\mathcal{C}_{-1}^{(m)}$,
the denominator should be interpreted as $n=1$.

\subsection{NLO ingredients}
\label{subsec:NLO-ingred}

At NLO, the required coefficients are $\mathcal{C}_1^{(1)}$ and $\mathcal{C}_0^{(1)}$,
given by $n=1$ and $n=0$ respectively in
\begin{equation}
\label{eq:Cn1}
\begin{aligned}
    \mathcal{C}_n^{(1)}&(\{p\},M,\mu_R)=
    f_a(x_a,\mu_F)\,f_b(x_b,\mu_F)\,|\vec{C}^{(0)}(\{p\},\mu_R)|^2
    \left[J_{c,n}^{(1)}(\lambda_c)
    +\widehat{S}^{(1)}_{\kappa,n}\!\left(\{\widehat{q}_\ell\},\lambda_s\right)\right.\\
    &\left.+\sum_{j}\int\frac{dz}{z}\,\mathcal{I}_{aj,n}^{(1)}(z,\lambda_F,\lambda_a)
    \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    +\sum_{j}\int\frac{dz}{z}\,\mathcal{I}_{bj,n}^{(1)}(z,\lambda_F,\lambda_b)
    \frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\right]\,.
\end{aligned}
\end{equation}
The beam function coefficients are related to the
coefficients $\mathcal{I}_{ij}$ by
\begin{equation}
  B_{i,n}^{(m)}(x_i,\mu_R,\mu_F,\lambda_i) =
  \sum_j\int\frac{dz}{z}\,\mathcal{I}_{ij,n}^{(m)}\!\left(z,\lambda_F,\lambda_i\right)
  f_j\!\left(\frac{x_i}{z},\mu_F\right)\,.
\end{equation}
We have substituted these directly into eq.~\eqref{eq:Cn1} for a
more transparent comparison with \textsc{ARES}, and refer to
$\mathcal{I}_{ij}$ as the beam function coefficients throughout.

The functions in Eq.~\eqref{eq:Cn1} are expanded in logarithms of the $\lambda$
parameters as
\begin{subequations}
\label{eq:subt-mn-exp}
\begin{align}
    J_{c,n}^{(m)}(\lambda_c) &= J_{c,n}^{(m)}
    + \sum_{r=1}^{2m-1-n}\frac{(n+r)!}{n!\,r!}\,J_{c,n+r}^{(m)}\,\ln^r\!\lambda_c\,, \\
    \widehat{S}_{\kappa,n}^{(m)}(\{\hat{q}_\ell\},\lambda_s) &=
    \widehat{S}_{\kappa,n}^{(m)}(\{\hat{q}_\ell\})
    + \sum_{r=1}^{2m-1-n}\frac{(n+r)!}{n!\,r!}\,
    \widehat{S}_{\kappa,n+r}^{(m)}(\{\hat{q}_\ell\})\,\ln^r\!\lambda_s\,, \\
    \mathcal{I}_{aj,n}^{(m)}(z,\lambda_F,\lambda_a) &=
    \mathcal{I}_{aj,n}^{(m)}(z,\lambda_F)
    + \sum_{r=1}^{2m-1-n}\frac{(n+r)!}{n!\,r!}\,
    \mathcal{I}_{aj,n+r}^{(m)}(z,\lambda_F)\,\ln^r\!\lambda_a\,,
\end{align}
\end{subequations}
and analogously for $\mathcal{I}_{bj,n}^{(m)}(z,\lambda_F,\lambda_b)$. 
The $\lambda$
parameters are defined as
\begin{equation}
    \lambda_c = \frac{Q_c M}{\mu_R^2}\,, \qquad
    \lambda_s = \frac{M}{\mu_R}\,, \qquad
    \lambda_a = \frac{Q_a M}{\mu_R}\,, \qquad
    \lambda_b = \frac{Q_b M}{\mu_R}\,, \qquad
    \lambda_F = \frac{\mu_R^2}{\mu_F^2}\,.
\end{equation}

We first consider the coefficient of the leading logarithm, $\mathcal{C}^{(1)}_1$.
Substituting the ingredients found in Eq.~\eqref{eq:C11-ingred}
into Eq.~\eqref{eq:Cn1} gives
\begin{equation}
    \mathcal{C}^{(1)}_1=-4\left(C_a+C_b+C_c\right)\,.
\end{equation}
Applying Eq.~\eqref{eq:coeff-convent} immediately yields
\begin{equation}
    h_{12}=-\left(C_a+C_b+C_c\right)\,,
\end{equation}
which is in agreement with the \textsc{ARES} coefficient shown in Eq.~\eqref{eq:h12}.

We next consider $\mathcal{C}^{(1)}_0$. Substituting terms found in Eq.~\eqref{eq:C10-ingred} into Eq.~\eqref{eq:Cn1} gives
\begin{equation}
\label{eq:C10-init}
    \begin{aligned}
    \mathcal{C}^{(1)}_0={}&-2\sum_{\ell} \gamma_{\ell}^{(0)}
    +\sum_{j}\int\frac{dz}{z}2P_{aj}^{(0)}(z)\frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    +\sum_{j}\int\frac{dz}{z}2P_{bj}^{(0)}(z)\frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\\
    &+4C_a\ln\frac{Q_a M}{\mu_R^2}+4C_b\ln\frac{Q_b M}{\mu_R^2}
    +4C_c\ln\frac{Q_c M}{\mu_R^2} \\
    &-4\left[(C_a-C_b-C_c)\ln\frac{Q_{bc}^2}{Q_b Q_c}
    +(C_b-C_a-C_c)\ln\frac{Q_{ac}^2}{Q_a Q_c}
    +(C_c-C_a-C_b)\ln\frac{Q_{ab}^2}{Q_a Q_b}\right]\\
    &-8\ln\frac{M}{\mu_R}\sum_{\ell}C_{\ell}\,,
    \end{aligned}
\end{equation}
where $\gamma^{(0)}_{\ell}$ are given in Eq.~\eqref{eq:gammal-flav} and $P_{aj}^{(0)}(z)$ in Eq.~\eqref{eq:full-AP}. The $\mu_R$ dependence cancels throughout, giving rise to
\begin{equation}
\label{eq:C01-simplified}
    \begin{aligned}
    \mathcal{C}^{(1)}_0={}&-2\sum_{\ell} \gamma_{\ell}^{(0)}
    +\sum_{j}\int\frac{dz}{z}2P_{aj}(z)\frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    +\sum_{j}\int\frac{dz}{z}2P_{bj}(z)\frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\\
    &+4C_a\ln\frac{Q_{ab}^2 Q_{ac}^2}{Q_{bc}^2 MQ_a}
    +4C_b\ln\frac{Q_{ab}^2 Q_{bc}^2}{Q_{ac}^2 MQ_b}
    +4C_c\ln\frac{Q_{ac}^2 Q_{bc}^2}{Q_{ab}^2 MQ_c}\,.
    \end{aligned}
\end{equation}
Applying Eq.~\eqref{eq:coeff-convent}, we obtain
\begin{equation}
\label{eq:SCET-h11}
    \begin{aligned}
    h_{11}={}&\sum_{\ell}\gamma^{(0)}_{\ell}
    -\sum_{j}\int\frac{dz}{z}P_{aj}(z)\frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    -\sum_{j}\int\frac{dz}{z}P_{bj}(z)\frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\\
    &-2C_a\ln\frac{Q_{ab}^2 Q_{ac}^2}{Q_{bc}^2 MQ_a}
    -2C_b\ln\frac{Q_{ab}^2 Q_{bc}^2}{Q_{ac}^2 MQ_b}
    -2C_c\ln\frac{Q_{ac}^2 Q_{bc}^2}{Q_{ab}^2 MQ_c}\,,
    \end{aligned}
\end{equation}
which agrees with the \textsc{ARES} coefficient shown in Eq.~\eqref{eq:h11}.

We turn our attention to the $\delta(\tau_1)$ coefficient
\begin{equation}
\begin{aligned}
    \mathcal{C}_{-1}^{(1)}&(\{p\},M,\mu_R)=
    f_a(x_1,\mu_F)\,f_b(x_2,\mu_F)
    \left(\vec{C}^{\dagger(1)}\vec{C}^{(0)}+\vec{C}^{\dagger(0)}\vec{C}^{(1)}\right)(\{p\},\mu_R) \\
    &+f_a(x_a,\mu_F)\,f_b(x_b,\mu_F)\,|\vec{C}^{(0)}(\{p\},\mu_R)|^2
    \left[J_{c,-1}^{(1)}(\lambda_c)
    +\widehat{S}^{(1)}_{\kappa,-1}\!\left(\{\widehat{q}_\ell\},\lambda_s\right)\right.\\
    &\left.+\sum_{j}\int\frac{dz}{z}\mathcal{I}_{aj,-1}^{(1)}(z,\lambda_F,\lambda_a)
    \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    +\sum_{j}\int\frac{dz}{z}\mathcal{I}_{bj,-1}^{(1)}(z,\lambda_F,\lambda_b)
    \frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\right]\,,
\end{aligned}
\end{equation}
where the ingredients are found in Eq.~\eqref{eq:Cd1-ingred} and
the terms are expanded as
\begin{subequations}
\begin{align}
    J_{c,-1}^{(m)}(\lambda_c) &=
    J_{c,-1}^{(m)}+\sum_{n=0}^{2m-1}J_{c,n}^{(m)}\,\frac{\ln^{n+1}\!\lambda_c}{n+1}\,, \\
    \widehat{S}_{\kappa,-1}^{(m)}(\lambda_s) &=
    \widehat{S}_{\kappa,-1}^{(m)}+\sum_{n=0}^{2m-1}\widehat{S}_{\kappa,n}^{(m)}\,
    \frac{\ln^{n+1}\!\lambda_s}{n+1}\,,\\
    \mathcal{I}_{aj,-1}^{(m)}(z,\lambda_F,\lambda_a) &=
    \mathcal{I}_{aj,-1}^{(m)}(z,\lambda_F)
    + \sum_{n=0}^{2m-1}\mathcal{I}_{aj,n}^{(m)}(z,\lambda_F)\,
    \frac{\ln^{n+1}\!\lambda_a}{n+1}\,,
\end{align}
\end{subequations}
and analogously for $\mathcal{I}_{bj,-1}^{(m)}$. 

The coefficient $h_{10}$ maps to $\mathcal{C}_{-1}^{(1)}/(-2)$ via
eq.~\eqref{eq:coeff-convent} with $m=1$, $n=1$. Unlike the
logarithmic coefficients validated in the following subsections,
$\mathcal{C}_{-1}^{(1)}$ receives a contribution from the hard
Wilson coefficient $\vec{C}^{(1)}$, which is process-specific and
must be computed explicitly. Since the $Z$ + jet production
virtual corrections are not  publicly available, we use the
results of $hh\to H $+ jet from Ref.~\cite{Ravindran:2002dc}.
The calculation of this for $q\bar{q}\to H g$
has been done in Appendix.~\ref{ap:H1-extraction}, we only do this 
for this channel and so the validation will follow for this channel. 
The remaining channels follow in the same way.

Using Eq.~\eqref{eq:C1-SCET}, along with ingredients from
Eq.~\eqref{eq:Cd1-ingred}, we obtain for the $q\bar{q}\to g H$
channel
\begin{equation}
	\begin{aligned}
	C^{(1)}_{-1}=\;&\frac{67\,C_A-10\,n_f}{9}
	+\frac{19\pi^2}{6}(2C_F-C_A)
	-\beta_0\ln\frac{M\,Q_c\,Q_{ab}^2}{\mu_R^6}\\[1ex]
	&+\sum_{j}\int\frac{dz}{z} \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}2I_{qj}^{(1)}+\sum_{j}\int\frac{dz}{z}\frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}2I_{\bar{q}j}^{(1)}\\
	&+\sum_{j}\int\frac{dz}{z}\left(2P_{qj}\ln\frac{Q_a\,M}{Q^2}+2P_{qj}\ln\frac{Q^2}{\mu_F^2}+3C_F\ln\frac{Q^2}{Q_a\,M}\right)
    \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}\\
    &+\sum_{j}\int\frac{dz}{z}\left(2P_{\bar{q}j}\ln\frac{Q_b\,M}{Q^2}+2P_{qj}\ln\frac{Q^2}{\mu_F^2}+3C_F\ln\frac{Q^2}{Q_b\,M}\right)
    \frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\\[1ex]
	&+3C_F\ln\frac{Q_{ab}}{Q^2}
	+3C_F\ln\frac{Q_{ab}}{Q^2}\\[1ex]
	&+2C_F\left[
	\ln^2\frac{M}{Q_a}
	+\ln^2\frac{M}{Q_b}
	-\ln^2\frac{Q_{ab}}{M^2}
	\right]
\\[1ex]
	&+C_A\left[
	2\ln^2\frac{M}{Q_c}
	+\ln^2\frac{M^2}{Q_{ab}}
	-\ln^2\frac{M^2}{Q_{ac}}
	-\ln^2\frac{M^2}{Q_{bc}}
	\right]\\[3ex]
	&+(C_A-2C_F)\left(
	\ln^2\frac{Q_{ab}}{Q_aQ_b}
	+2(I_{ab,c}+I_{ba,c})
	\right)\\[1ex]
	&-C_A\left(
	\ln^2\frac{Q_{ac}}{Q_aQ_c}
	+2(I_{ac,b}+I_{ca,b})
	\right)\\[1ex]
	&-C_A\left(
	\ln^2\frac{Q_{bc}}{Q_bQ_c}
	+2(I_{bc,a}+I_{cb,a})
	\right)\\[2ex]
	&+C_A\left[
	\frac{152}{9}
	+4\operatorname{Li}_2\left(\frac{Q_{ab}-M^2}{Q_{ab}}\right)
	+2\ln\frac{Q_{ab}}{Q_{ac}}
	\ln\frac{Q_{ab}}{Q_{bc}}
	\right]\\[1ex]
	&+C_F\left(
	-16
	+4\operatorname{Li}_2\left(\frac{-Q_{ac}}{M^2}\right)
	+4\operatorname{Li}_2\left(\frac{-Q_{bc}}{M^2}\right)
	-2\ln^2\frac{Q_{ac}+M^2}{Q_{ac}}
	-2\ln^2\frac{Q_{bc}+M^2}{Q_{bc}}
\right.
\\[1ex]
	&\left.
	+2\ln^2\frac{M^2+Q_{ac}}{M^2}
	+2\ln^2\frac{M^2+Q_{bc}}{M^2}
	+2\ln^2\frac{Q_{ac}}{Q_{bc}}
	\right)
	-\frac{20}{9}n_f\,.
	\end{aligned}
\end{equation}

The \textsc{ARES} coefficient $h_{10}$ for the $q\bar q\to g H$ channel is
\begin{equation}
\begin{aligned}
	h_{10}=\;&\frac{67\,C_A-10\,n_f}{18}
	+\frac{19\pi^2}{12}(2C_F-C_A)
	-\beta_0\ln\frac{M\,Q_c\,Q_{ab}^2}{\mu_R^6}\\[2ex]
	&+\sum_{j}\int\frac{dz}{z} \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}2I_{qj}^{(1)}+\sum_{j}\int\frac{dz}{z}\frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}2I_{\bar{q}j}^{(1)}\\
	&+\sum_{j}\int\frac{dz}{z}\left(P_{qj}(z)\ln\frac{Q_a\,M}{Q^2}+P_{qj}(z)\ln\frac{Q^2}{\mu_F^2}+\frac{3}{2}C_F\ln\frac{Q^2}{Q_a\,M}\right)
    \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}\\
    &+\sum_{j}\int\frac{dz}{z}\left(P_{\bar{q}j}(z)\ln\frac{Q_b\,M}{Q^2}+P_{qj}(z)\ln\frac{Q^2}{\mu_F^2}+\frac{3}{2}C_F\ln\frac{Q^2}{Q_b\,M}\right)
    \frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\\[1ex]
	&+\frac{3}{2}C_F\ln\frac{Q_{ab}}{Q^2}
	+\frac{3}{2}C_F\ln\frac{Q_{ab}}{Q^2}\\[1ex]
	&+C_F\left[
	\ln^2\frac{M}{Q_a}
	+\ln^2\frac{M}{Q_b}
	-\ln^2\frac{Q_{ab}}{M^2}
	\right]
\\[1ex]
	&+\frac{C_A}{2}\left[
	2\ln^2\frac{M}{Q_c}
	+\ln^2\frac{M^2}{Q_{ab}}
	-\ln^2\frac{M^2}{Q_{ac}}
	-\ln^2\frac{M^2}{Q_{bc}}
	\right]\\[2ex]
	&+\frac{(C_A-2C_F)}{2}
	\left(
	\ln^2\frac{Q_{ab}}{Q_aQ_b}
	+2(I_{ab,c}+I_{ba,c})
	\right)\\[1ex]
	&-\frac{C_A}{2}
	\left(
	\ln^2\frac{Q_{ac}}{Q_aQ_c}
	+2(I_{ac,b}+I_{ca,b})
	\right)\\[1ex]
	&-\frac{C_A}{2}
	\left(
	\ln^2\frac{Q_{bc}}{Q_bQ_c}
	+2(I_{bc,a}+I_{cb,a})
	\right)\\
	&+\frac{C_A}{2}\left[
	\frac{152}{9}
	+4\operatorname{Li}_2\left(\frac{Q_{ab}-M^2}{Q_{ab}}\right)
	+2\ln\frac{Q_{ab}}{Q_{ac}}
	\ln\frac{Q_{ab}}{Q_{bc}}
	\right]\\[1ex]
	&+\frac{C_F}{2}\left(
	-16
	+4\operatorname{Li}_2\left(\frac{-Q_{ac}}{M^2}\right)
	+4\operatorname{Li}_2\left(\frac{-Q_{bc}}{M^2}\right)
	-2\ln^2\frac{Q_{ac}+M^2}{Q_{ac}}
	-2\ln^2\frac{Q_{bc}+M^2}{Q_{bc}}
\right.
\\[1ex]
	&\left.
	+2\ln^2\frac{M^2+Q_{ac}}{M^2}
	+2\ln^2\frac{M^2+Q_{bc}}{M^2}
	+2\ln^2\frac{Q_{ac}}{Q_{bc}}
	\right)
	-\frac{20}{18}n_f\,,
\end{aligned}
\end{equation}
where we have used the relation
\begin{equation}
\label{eq:I1-ARES}
I_{ij}^{(1)}(z) =
-\hat{P}_{ij}^{(\varepsilon)}(z)
+\hat{P}_{ij}(z;0)\!\left[
\left(\frac{\ln(1-z)}{1-z}\right)_{\!\!+}(1-z)
-\ln z\right]
+\frac{\pi^2}{6}\,\delta_{ij}\,C_i\,\delta(1-z)\,,
\end{equation}
to express the SCET beam function scalar kernels $I_{qj}^{(1)}$ and
$I_{\bar{q}j}^{(1)}$ in terms of \textsc{ARES} ingredients: the
first two terms correspond to part $C_{\mathrm{hc},\ell}^{(1)}$
and $\delta\mathcal{F}_{\rm rec}$ for ISR, shown in Eqs.~\eqref{eq:Cij-ISR}
and~\eqref{eq:dFrec-ISR}.

Overall, one can see that this gives rise to $\mathcal{C}^{(1)}_{-1}=2h_{10}$
as given in Eq.~\eqref{eq:coeff-convent}.







\subsection{NNLO ingredients}
\label{subsec:NNLO-ingred}

Unlike the NLO case, there is no compact master expression for the
NNLO subtraction coefficients. For brevity the scale arguments are suppressed throughout this subsection. The contributions which make up these NNLO coefficients are organised into three types of quantity. The genuine $m$-loop contributions from the jet, beam, and soft functions as
\begin{equation}
\begin{aligned}
\label{eq:Xn-def}
    X_n^{(m)} \equiv &f_a(x_a,\mu_F)\,f_b(x_b,\mu_F)\,|\vec{C}^{(0)}|^2
    \left[J_{c,n}^{(m)} + \widehat{S}_{\kappa,n}^{(m)} \right.\\
&  \left.  + \sum_j\int\frac{dz}{z}\,\mathcal{I}_{aj,n}^{(m)}\frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    + \sum_j\int\frac{dz}{z}\,\mathcal{I}_{bj,n}^{(m)}\frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\right]\,.
    \end{aligned}
\end{equation}
The cross terms involving the one-loop
virtual amplitude $\vec{C}^{(1)}$ are
\begin{equation}
\label{eq:Xn11-def}
\begin{aligned}
    X_n^{(1+1)} \equiv& f_a(x_a,\mu_F)\,f_b(x_b,\mu_F)
    \!\left(\vec{C}^{\dagger(0)}\vec{C}^{(1)}+\vec{C}^{\dagger(1)}\vec{C}^{(0)}\right)
    \left[J_{c,n}^{(1)} + \widehat{S}_{\kappa,n}^{(1)} \right.\\
  &\left.  + \sum_j\int\frac{dz}{z}\,\mathcal{I}_{aj,n}^{(1)}\frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    + \sum_j\int\frac{dz}{z}\,\mathcal{I}_{bj,n}^{(1)}\frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\right]\,.
    \end{aligned}
\end{equation}
Finally, the cross-terms between two one-loop functions arising from the
$\mathcal{L}_n\otimes\mathcal{L}_m$ convolution are
\begin{equation}
\label{eq:Xnm11-def}
\begin{aligned}
    X_{n,m}^{(1+1)} \equiv{}&
    f_a(x_a,\mu_F)\,f_b(x_b,\mu_F)\,|\vec{C}^{(0)}|^2
    \left[J_{c,n}^{(1)}\,\widehat{S}_{\kappa,m}^{(1)}\right.\\
    &+\!\left(J_{c,m}^{(1)}+\widehat{S}_{\kappa,m}^{(1)}\right)\!
    \left(\sum_j\int\frac{dz}{z}\,\mathcal{I}_{aj,n}^{(1)}
    \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    +\sum_j\int\frac{dz}{z}\,\mathcal{I}_{bj,n}^{(1)}
    \frac{f_j(x_b/z,\mu_F)}{f_b(x_b,\mu_F)}\right)\\
    &\left.+\sum_j\int\frac{dz}{z}\,\mathcal{I}_{aj,n}^{(1)}
    \frac{f_j(x_a/z,\mu_F)}{f_a(x_a,\mu_F)}
    \cdot\sum_{j'}\int\frac{dz'}{z'}\,\mathcal{I}_{bj',m}^{(1)}
    \frac{f_{j'}(x_b/z',\mu_F)}{f_b(x_b,\mu_F)}\right]\,.
\end{aligned}
\end{equation}
With these definitions, the NNLO subtraction coefficients are
\begin{subequations}
\label{eq:NNLO-Cn}
\begin{align}
    \mathcal{C}_3^{(2)} &= X_3^{(2)} + X_{1,1}^{(1+1)}\,, 
    \label{eq:C32}\\
    \mathcal{C}_2^{(2)} &= X_2^{(2)} + \frac{3}{2}\!\left(
        X_{0,1}^{(1+1)} + X_{1,0}^{(1+1)}\right)\,,
    \label{eq:C22}\\
    \mathcal{C}_1^{(2)} &= X_1^{(2)} + X_1^{(1+1)} + 2X_{0,0}^{(1+1)}
        - \frac{\pi^2}{3}X_{1,1}^{(1+1)}+X_{1,-1}^{(1+1)}+X_{-1,1}^{(1+1)}\,,
    \label{eq:C12}\\
    \mathcal{C}_0^{(2)} &= X_0^{(2)} + X_0^{(1+1)}
        - \frac{\pi^2}{6}\!\left(X_{0,1}^{(1+1)}+X_{1,0}^{(1+1)}\right)
        + 2\zeta_3\,X_{1,1}^{(1+1)}+X_{0,-1}^{(1+1)}+X_{-1,0}^{(1+1)}\,.
    \label{eq:C02}
\end{align}
\end{subequations}
The last two terms Eqs.~\eqref{eq:C12} and~\eqref{eq:C02} are not present in Ref.~\cite{Gaunt:2015pea}. Please note that
we believe this to be an error on their side and these terms must contribute to make
up the full coefficient. \ABquestion{Need to make sure this is correct and not
something different about their subtraction scheme?}

Starting from $\mathcal{C}^{(2)}_3$, we find that by
substituting the ingredients from Eq.~\eqref{eq:C23-ingred} into
Eq.~\eqref{eq:C32} we obtain
\begin{equation}
    \mathcal{C}^{(2)}_3=8\left(C_a+C_b+C_c\right)^2\,.
\end{equation}
which via Eq.~\eqref{eq:coeff-convent} gives
\begin{equation}
    h_{24}=\frac{1}{2}\left(\sum_{\ell}C_a+C_b+C_c\right)^2\,,
\end{equation}
in agreement with the \textsc{ARES} coefficient shown in Eq.~\eqref{eq:h24}.

Next, the coefficient $\mathcal{C}^{(2)}_2$, is found by
substituting the terms from Eq.~\eqref{eq:C22-ingred} into
Eq.~\eqref{eq:C22} 
\begin{equation}
    \mathcal{C}_{2}^{(2)} = 24\pi\beta_0\left(C_a+C_b+C_c\right)
    +12\left(-\frac{\mathcal{C}^{(1)}_{0}}{2}\right)\left(C_a+C_b+C_c\right)\,.
\end{equation}
Applying Eq.~\eqref{eq:coeff-convent} gives
\begin{equation}
    h_{23}=-2\pi\beta_{0}\left(C_a+C_b+C_c\right)-h_{11}\left(C_a+C_b+C_c\right)\,,
\end{equation}
which agrees exactly with the \textsc{ARES} coefficient shown in Eq.~\eqref{eq:h23}.

The remaining coefficients $\mathcal{C}^{(2)}_1$ and
$\mathcal{C}^{(2)}_0$ are considerably more involved, depending on
significantly more complex ingredients as shown in Eqs.~\eqref{eq:C21-ingred} and~\eqref{eq:C20-ingred}. We therefore validate $h_{22}$
and $h_{21}$ symbolically using \textsc{Mathematica}. Snippets of this code
are presented in Appendix~\ref{ap:symb-val}, while the actual code may be found in 
Ref.~\cite{peake2026validation}.

In the next section, we provide an alternative analytic 
validation for the ISR procedure as a whole. 

\section{Validation: lund-tree observables}
\label{sec:LTS-val}

In this section we validate the ISR formalism developed in
Sec.~\ref{sec:NNLL-initial} by comparing it with the NNLL resummation of Ref.~\cite{vanBeekveld:2025zjh}, which derives results for a class
of observables known as Lund-tree Shapes (LTS) in hadron--hadron, DIS,
and $e^+e^-$ collisions. For hadron--hadron collisions, the NNLL
resummation is restricted to colour-singlet production with no resolved
jets ($N=0$), providing a clean setting in which to compare
initial-state hard-collinear corrections.

Restricting the \textsc{ARES} master formula Eq.~\eqref{eq:Sigma-NNLL}
 to the same $N=0$ configuration leaves only the two incoming legs 
 $\ell=1,2$, such that the only Born channel is $q\bar q\to Z$. The NNLL cumulant becomes
\begin{multline}
  \label{eq:Sigma-qqbar}
  \Sigma_{q\bar q}(\{p\},v)= e^{-R_{\rm NNLL}(v)} \frac{f_{q}(x_1,v^{\frac{1}{a+b_1}}Q)}{f_{q}(x_1,\mu_F)} \frac{f_{\bar q}(x_2,v^{\frac{1}{a+b_2}}Q)}{f_{\bar q}(x_2,\mu_F)}
  \times \\ \times
  \left(\mathcal{F}^{(q \bar q)}_{\rm NLL}(\lambda)\left(1+\frac{\alpha_s(Q)}{2\pi}H_{q \bar q}^{(1)}(Q,\mu_R)
      +\frac{\alpha_s(Q v^{\frac{1}{a+b_\ell}})}{2\pi}\sum_{\ell}C_{\mathrm{hc},\ell}^{(1)}\right) +\frac{\alpha_s(Q)}{\pi} \delta\mathcal{F}^{(q\bar q)}_{\rm NNLL}(\lambda)\right)\,,
\end{multline}
with 
\begin{equation}
\label{eq:dFNNLL-qq}
	 \delta\mathcal{F}^{(q\bar q)}_{\rm NNLL} = \delta\mathcal{F}_{\rm sc}+\delta\mathcal{F}_{\rm hc}+\delta\mathcal{F}_{\rm wa}+\delta\mathcal{F}_{\rm correl}+\delta \mathcal{F}_{\mathrm{rec}}\,.
\end{equation}
Note that $\Delta\mathcal{F}_{\rm wa}$ and
$\Delta\mathcal{F}_{\rm rec}$ are absent from Eq.~\eqref{eq:dFNNLL}, 
since these corrections were first introduced 
for three-jet events in Ref.~\cite{Arpino:2019ozn} and
require a final-state coloured leg, which is absent in the $N=0$ 
configuration. This means that the $\Delta\mathcal{F}_{\rm rec}$
ISR contribution cannot be tested against the LTS comparison, and
we do not attempt to validate it here.

We restrict ourselves to the additive LTS observable $S_b^{(0)}$,
for which the NNLL correction contains a comparable ISR
term. \ABquestion{Is this why we chose this observable? Also are we
meant to say they aren't event shapes?} This allows a term-by-term
comparison of the ISR contributions. The observable is
\begin{equation} 
\label{eq:Sb0-def} 
S_b^{(0)} = \sum_{j\in\mathcal{I}_0}v_b^j \,, \qquad v_b^{j_i} = \frac{p_{t,i}}{M}e^{-b|y_i|}\,,
\end{equation} 
where $(p_{t,i},y_i)$ are the transverse momentum and rapidity of
jet $i$ with respect to the beam axis, $M$ is the invariant mass of
the colour singlet (taken as the hard scale), and
$\mathcal{I}_0\equiv\{j_1,j_2,\dots,j_m\}$ is the list of jets from
the C/A algorithm of radius $R$.

Setting $Q=M$ fixes the single-emission parameters to
\begin{equation}
\label{eq:LTS-1-emsn}
	a=1, \quad b_\ell=b, \quad d_\ell=1, \quad g_\ell(\phi)=1,\quad \ell=1,2\,.
\end{equation}
Throughout this section we identify $x_q\equiv x_1$ and
$x_{\bar{q}}\equiv x_2$, where $x_\ell$ is the momentum fraction of
the parton on leg $\ell$ as defined in Sec.~\ref{sec:setup}. \ABquestion{I am
really unsure about the notation between the two formalisms -- I
am concerned it is quite confusing.}

In our notation, the NNLL cumulant of
Ref.~\cite{vanBeekveld:2025zjh} with $Q=M$ is \ABquestion{Do we need to explain
how we get this? I don't quite understand how we get this cumulant 
from their equations?}
\begin{equation}
\label{eq:LTS-cum}
\begin{aligned}
\Sigma^{pp}&(\{p\},v)=e^{-R_{\mathrm{NNLL}}(v)}
\frac{f_q(x_q,\mu_F v^{\frac{1}{1+b}})f_{\bar q}(x_{\bar q},\mu_F v^{\frac{1}{1+b}})}
{f_q(x_q,\mu_F )f_{\bar q}(x_{\bar q},\mu_F )}
\times\\
&\left[\mathcal{F}_{\mathrm{NLL}}(v)\left(
1+\frac{\alpha_s(\mu_R)}{2\pi}H_{q\bar q}^{(1)}(\{p\})+
\right.\right.\\
&\left. \left.+\frac{\alpha_s(\mu_R)}{2\pi}\frac{1+b}{1+b-2\lambda}
\sum_j\left(
\int_{x_q}^1\frac{dz}{z}C_{qj}^{(1), \rm LTS}(z)
\frac{f_j\left(x_q/z,\mu_F v^{\frac{1}{1+b}}\right)}{f_{ q}(x_{q},\mu_F v^{\frac{1}{1+b}})}
+\{q\leftrightarrow \bar q\}
\right)
\right)+\right.\\
&\left.+\frac{\alpha_s(\mu_R)}{\pi}\delta\mathcal{F}^{ pp}_{\mathrm{NNLL}}(v)\right]+\delta\Sigma_{\mathrm{DGLAP}}^{pp}(v)\,,
\end{aligned}
\end{equation}
where 
\begin{equation}
\label{eq:NNLL-LTS}
	\delta\mathcal{F}^{pp}_{\mathrm{NNLL}}(v)=\delta\mathcal{F}^{pp}_{\mathrm{sc}}(v)+\delta\mathcal{F}^{pp}_{\mathrm{wa}}(v)+\delta\mathcal{F}^{pp}_{\mathrm{hc}}(v)+\delta\mathcal{F}^{pp}_{\mathrm{correl}}(v)+\delta\mathcal{F}^{pp}_{\mathrm{clust}}(v)\,,
\end{equation}
and
\begin{equation}
\label{eq:DGLAP-add}
\begin{aligned}
	\delta\Sigma_{\mathrm{DGLAP}}^{pp}(v)&=e^{-R_{\mathrm{NNLL}}(v)}\frac{\alpha_s(\mu_R)}{\pi}\frac{1}{1+b-2\lambda}\frac{d\FNLL}{d\RpNLL}\times\\
	&\times\sum_j\left(
\int_{x_q}^1\frac{dz}{z}P_{qj}(z)
\frac{f_j\left(x_q/z,\mu_F v^{1/(1+b)}\right)}{f_{ q}(x_{q},\mu_F v^{1/(1+b)})}
+\{q\leftrightarrow \bar q\}
\right)\,,
\end{aligned}
\end{equation}
where $P_{qi}(z)$ are the full AP splitting functions defined in Eq.~\eqref{eq:full-AP}. 

The two cumulants have a very similar structure. Their correspondence
can be summarised as follows.

The radiator $R_{\rm NNLL}$, hard coefficient $H^{(1)}_{q\bar{q}}$,
and NLL function $\mathcal{F}_{\rm NLL}$ are identical in the two
formulations. The NNLL corrections of soft origin ---
$\delta\mathcal{F}_{\rm sc}$, $\delta\mathcal{F}_{\rm wa}$, and
$\delta\mathcal{F}_{\rm correl}$ --- are also identical, since they
depend only on soft-emission kinematics and do not modify the
longitudinal momentum fractions of the incoming partons. As
Ref.~\cite{vanBeekveld:2025zjh} uses the same \textsc{ARES} treatment
for these contributions, they require no further validation here. The
additional correction $\delta\mathcal{F}^{pp}_{\rm clust}$ in
eq.~\eqref{eq:NNLL-LTS} arises from the jet-clustering dependence of
$S_b^{(0)}$ and has no counterpart in \textsc{ARES}, which measures
hadrons directly (corresponding to infinite jet radius).

The non-trivial comparison concerns the hard-collinear contributions.
In \textsc{ARES} these are organised into $C_{{\rm hc},\ell}^{(1)}$,
$\delta\mathcal{F}_{\rm hc}$, and $\delta\mathcal{F}_{\rm rec}$,
whereas Ref.~\cite{vanBeekveld:2025zjh} introduces
$C_{qi}^{(1),\rm LTS}$, $\delta\mathcal{F}^{pp}_{\rm hc}$, and
$\delta\Sigma^{pp}_{\rm DGLAP}$. Although the organisation differs,
we show below that the two formulations contain identical ISR
corrections.

The coefficient function of Ref.~\cite{vanBeekveld:2025zjh}
is\footnote{In Ref.~\cite{vanBeekveld:2025zjh} the final term is
written as $\ln(M^2/\mu_F^2)\hat{P}_{ij}(z;0)$, which appears to be
an error since it introduces an uncancelled collinear singularity --
this has been confirmed with the authors of the paper.}
\begin{multline}
  \label{eq:CkiLTS}
  \sum_j\int_{x_q}^1\frac{dz}{z}\,C_{qj}^{(1),\rm LTS}(z)
  \frac{f_j(x_q/z,\mu_F v^{\frac{1}{1+b}})}
  {f_q(x_q,\mu_F v^{\frac{1}{1+b}})}   = \sum_j\int_{x_q}^1\frac{dz}{z}\,
  \frac{f_j(x_q/z,\mu_F v^{\frac{1}{1+b}})}
  {f_q(x_q,\mu_F v^{\frac{1}{1+b}})}
  \left[-\hat{P}_{qj}^{(\varepsilon)}(z)\right.\\
  \left.+\frac{2b}{1+b}\!\left(\!
  \left(\frac{\ln(1-z)}{1-z}\right)_+
  -\frac{\ln z}{1-z}\right)
  (1-z)\hat{P}_{qj}(z;0)
  +\ln\frac{M^2}{\mu_F^2}P_{qj}(z)\right]\,.
\end{multline}
The first two terms correspond to contributions that are treated
separately in \textsc{ARES}, namely the hard-collinear coefficient
$C_{{\rm hc},\ell}^{(1)}$ and the recoil correction
$\delta\mathcal{F}_{\rm rec}$. To see this, we
specialise Eq.~\eqref{eq:Cell-ISR} to the parameters in
Eq.~\eqref{eq:LTS-1-emsn} with $Q=M$ and $E_\ell=M/2$:
\begin{multline}
  \label{eq:Cij-LTS-comp}
  \frac{\alpha_s(Mv^{\frac{1}{1+b}})}{2\pi}\sum_{\ell=1}^2
  C_{{\rm hc},\ell}^{(1)} =\\
  =\sum_j\!\left[\int_{x_1}^1\frac{dz}{z}
  \frac{f_j(x_1/z,Mv^{\frac{1}{1+b}})}
  {f_q(x_1,Mv^{\frac{1}{1+b}})}
  \!\left(-\hat{P}_{qj}^{(\varepsilon)}(z)
  +\frac{2b}{1+b}
  \!\left(\frac{\ln(1-z)}{1-z}\right)_+\!(1-z)\hat{P}_{qj}(z;0)
  \right)+\{q\leftrightarrow\bar{q}\}\right]\,.
\end{multline}
Using Eq.~\eqref{eq:dFrec-ISR} for additive observables with the
parameters of Eq.~\eqref{eq:LTS-1-emsn},
\begin{equation}
  \label{eq:dFrec-LTS-comp}
  \frac{\alpha_s(M)}{\pi}\delta\mathcal{F}_{\rm rec}(\lambda) =
  \frac{\alpha_s(Mv^{1/(1+b)})}{2\pi}
  \sum_j\!\left[\int_{x_1}^1\frac{dz}{z}
  \frac{f_j(x_1/z,\mu_F v^{\frac{1}{1+b}})}
  {f_q(x_1,\mu_F v^{\frac{1}{1+b}})}
  \!\left(-\frac{2b}{1+b}\hat{P}_{qj}(z;0)\ln z\right)
  +\{q\leftrightarrow\bar{q}\}\right]\,.
\end{equation}
Combining Eqs.~\eqref{eq:Cij-LTS-comp} and~\eqref{eq:dFrec-LTS-comp}
reproduces the first two terms of Eq.~\eqref{eq:CkiLTS} exactly. The
third term, $\ln(M^2/\mu_F^2)P_{qj}(z)$, arises because the PDFs are 
evaluated at different scales in the two formulations:
\textsc{ARES} uses $Mv^{1/(1+b)}$, while Ref.~\cite{vanBeekveld:2025zjh} uses $\mu_Fv^{1/(1+b)}$. The logarithm $\ln(M^2/\mu_F^2)$ is the 
DGLAP evolution between these two scales and compensates for the
difference between the two. The prefactor $(1+b)/(1+b-2\lambda)$ in Eq.~\eqref{eq:LTS-cum} arises from expressing the running coupling $\alpha_s(Mv^{1/(1+b)})$ in terms of $\lambda=\alpha_s(\mu_R)\beta_0\ln(1/v)$ at one loop. \ABquestion{Is this statement true about the $(1+b)/(1+b-2\lambda)$ factor arising? Really it should be a term of $\ln M/\mu_R$ as they have
$\alpha_s(\mu_R)$ but maybe this is neglected at NNLL?}

The remaining hard-collinear correction in the LTS formulation is
\begin{equation}
\label{eq:dFhc-LTS}
\delta\mathcal{F}^{pp}_{\mathrm{hc}}(v)
=-\frac{d\mathcal{F}_{\rm NLL}}{dR_{\rm NLL}}
\frac{1}{1+b-2\lambda}
\sum_{\ell}\gamma_\ell^{(0)}\,.
\end{equation}
Combining this with $\delta\Sigma^{pp}_{\rm DGLAP}$ in Eq.~\eqref{eq:DGLAP-add}, we find that at NNLL this gives rise to:
\begin{multline}
	\frac{\alpha_s(\mu_R)}{\pi}\delta\mathcal{F}^{pp}_{\mathrm{hc}}(v)+
	\delta\Sigma^{pp}_{\rm DGLAP}\simeq\\
	\frac{\alpha_s(\mu_R)}{\pi(1+b)}\frac{1+b}{1+b-2\lambda}\frac{d\mathcal{F}_{\rm NLL}}{dR_{\rm NLL}}\left[\int_{x_q}^1\frac{dz}{z}\left(P_{qj}(z)-\gamma^{(0)}_1\delta_{qj}\right)
\frac{f_j\left(x_q/z,\mu_F v^{1/(1+b)}\right)}{f_{ q}(x_{q},\mu_F v^{1/(1+b)})}
+\{q\leftrightarrow \bar q\}
\right]\,.
\end{multline}

This is equivalent to $\delta\mathcal{F}_{\rm hc}$ in \textsc{ARES}, and can see this explicitly by calculating
Eqs.~\eqref{eq:dFhc-ISR} for additive observables,
\begin{equation}
	\frac{\alpha_s(Q)}{\pi}\delta\mathcal{F}_{\rm hc}=\frac{\alpha_s(Qv^{1/(1+b)})}{\pi(1+b)}\frac{d\FNLL}{d\RpNLL}\sum_j \left[\int_{x_1}^1 \frac{dz}{z} \left(P_{qj}(z)-\gamma_1^{(0)}\delta_{qj}\right) \frac{f_j(x_{1}/z,Qv^{1/(a+b_\ell)})}{f_q(x_1,Qv^{1/(a+b_\ell)})} +\{q\leftrightarrow\bar{q}\}\right]\,.
\end{equation}
This is exactly the combination of the DLGAP term with $\delta\mathcal{F}_{\rm hc}^{pp}$. Consequently, the
additional DGLAP term introduced in
Ref.~\cite{vanBeekveld:2025zjh} does not represent a new physical effect, but rather a different organisation of the hard-collinear corrections. Once all terms are combined, the two resummation formalisms are equivalent. \ABquestion{I am 
concerned what I am saying is a bit too strong here?}

Throughout we identify $x_q\equiv x_1$ and $x_{\bar{q}}\equiv x_2$,
where $x_\ell$ denotes the momentum fraction of the parton on leg
$\ell$ as defined in Sec.~\ref{sec:setup}.


\section{Conclusion}
\todo{Answer these questions:}
\begin{enumerate}
\item What did we do in this project? \\
Showed a formalism for ISR/hadronic collisions 
with ARES.\\
The formalism is really the main result
for this piece of work, since the results from
one-jettiness have been discovered before. \\
Most notable thing we have found is that
the same cumulant holds fro both ISR and
FSR.
\item How could the project be extended? \\
To other observables, namely another additive observable, the 
transverse thrust, and a non-additive observable, the thrust minor.
Discuss how these would be different for different SCET based observables
i..\ SCET 1/2. \\ 
Go up to NNNLL. \\
Look at $H+$ jet process. \\
Extend to DIS process (although not really 
an extension as just 1 leg). \\
More jets is complicated as from 4 legs onwards the wide angle corrections become matrices in colour space
which is already a huge amount more work.
\item How can these results be used? \\
When matching to fixed order we can then use these results
to compare to experiment at the LHC. I believe the scientific community already have
the FO up to quite a high order therefore we need the resummed ones. \\
Used for non perturbative corrections? \\
Tests for $\alpha_s$ extraction - I think proton collisions
are the best for these? 
\end{enumerate}




